### 2024-12-02 paper/paper.qmd
Updated paper to include data section

---
title: "Analyzing the Pakistan Cricket Team's Performance (2002-2024)"
subtitle: "My subtitle if needed"
author: 
  - Muhammad Abdullah Motasim
thanks: "Code and data are available at: [https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data](https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
table-of-contents: true
toc-depth: 2
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(ggplot2)
library(arrow)
library(patchwork)

pakistan_match_data <- read_parquet(here("data/02-analysis_data/cleaned_pakistan_match_data.parquet"))
```

# Introduction
Pakistan’s cricket team holds a storied legacy, marked by remarkable highs and challenging lows. From their 1992 World Cup triumph to fluctuations in form over the decades, analyzing trends in performance offers insights into their journey and areas for strategic improvement.

Despite an extensive focus on player statistics and match highlights, there remains a lack of systematic statistical analysis to evaluate long-term performance trends. This paper fills this gap by analyzing data from 2002 to 2024 to understand what factors contribute to Pakistan’s success in international cricket.

We employ a linear regression model to analyze match outcomes, considering variables like year, opponent, and toss outcomes. The focus is on identifying trends and predicting the probability of victory under varying conditions. Our findings not only contribute to the literature on sports analytics but also provide actionable insights for coaches, analysts, and cricket administrators not only in Pakistan, but around the world.

The structure of this paper is as follows: @sec-data discusses the data types included in the raw data, the cleaning process for the data, and the reason for selecting the data set we did. @sec-model discusses model specification and justification for a Linear Regression model. @sec-results analyzes the trends and correlations between different variables utilizing tabular and graphical means. @sec-discussion discusses the results of @sec-results going into detail on what the simulation results can tell us about Pakistan cricket performance, as well as assumptions and limitations in data.

## Estimand
The estimand is the probability that Pakistan wins a cricket match, given the opponent, toss outcome, and year.

# Data {#sec-data}

The data for this study were obtained from the cricketdata R package [@cricketdata] and all analysis was performed using `R` [@citeR] alongside the following packages: `tibble` [@tibble], `readr` [@readr], `arrow` [@arrow], `tidyverse` [@tidyverse], `dplyr` [@dplyr], `testthat` [@testthat], `caret` [@caret], `rstanarm` [@rstanarm], `tidyr` [@tidyr], patchwork [@patchwork], and ggplot2 [@ggplot2].

As mentioned before the raw data was collected from the cricketdata library which contains data on international and other major cricket matches sourced from ESPN Cricinfo [@espncricinfo] and Cricsheet [@cricsheet]. It contains details on the individual match year, teams, result, and toss winner. Additionally, the data set features comprehensive career statistics for individual players, including total runs, total balls faced, total wickets, and more. This data set was chosen for its ease of implementation within R, as it is available as a package, and for its ability to scrape data from multiple reputable cricket sources, such as ESPN, thereby enhancing the overall reliability of the data set.

## Raw Data
The cricketdata library offers a variety of functions that provide data sets on different cricket statistics. However, for our analysis, we focus on two key areas: individual match statistics, to determine which matches Pakistan played in and how they performed, and player career statistics, to assess how the quality of players has evolved over time and its impact on Pakistan's overall performance. To extract the relevant data, we primarily used two functions from the library: fetch_cricinfo, which retrieves individual career performance data, and fetch_cricsheet, which provides detailed match information.

## Data Cleaning
The raw match data initially contained 820 observations across 25 variables, each detailing a match between two countries. The data cleaning process involved selecting only matches in which Pakistan played, and retaining only the variables relevant to our analysis. These variables include:

- `Team1` - The first team that participated in the match.
- `Team2` - The second team that participated in the match.
- `Date` - The date of the match (eg. 2008-01-02)
- `Winner` - The team that won the match between Team1 and Team2 (NA if the match was a draw).
- `Winner_wickets` - The number of wickets the winning team took (NA if the match was a draw).
- `Winner_runs` - The number of runs the winning team scored (NA if the match was a draw).
- `Toss_winner` - The team that won the coin toss.
- `Toss_decision` - The decision made by the toss winner, indicating whether they chose to field or bowl.

The raw player career data was broken into batting, bowling, and fielding categories with some players appearing in multiple data sets, thus the cleaning process involved combining the 3 data sets and ensuring proper handling of one player in each set. After this, only Pakistani cricket players were selected and variables which were important to use were retained. These variables are:

- `Player` - Name of the player.
- `Start` - The year in which the player began their career
- `End` - The year in which the player's career concluded.
- `Matches` - The total number of matches the player participated in throughout their career.
- `Bat_innings` -  The total number of batting innings the player participated in.
- `Bowl_innings` -  The total number of bowling innings the player participated in.
- `Field_innings` -  The total number of fielding innings the player participated in.
- `Bat_runs` - The total number of runs the player scored while batting.
- `Not_puts` - The number of times the player remained not out while batting.
- `Bat_average` - The player's batting average, calculated as the total number of runs scored divided by the number of times they were dismissed.
- `Bowl-runs` - The total number of runs conceded by the player while bowling. This statistic reflects how many runs the player allowed the opposition to score during their bowling spells.
- `Wickets` - The total number of wickets taken by the player in their career.
- `Economy` - The average number of runs the player concedes per over bowled. It is calculated by dividing the total number of runs given by the total number of overs bowled. A lower economy rate typically indicates a bowler who is more effective at restricting the opposition's scoring.
- `Dismissals` - The total number of dismissals the player has contributed to as a fielder or wicketkeeper.

After applying these cleaning procedures, we were left with 151 observations across 7 variables for the matches Pakistan has played in, and 344 observations across 14 variables for the career information of individual Pakistani cricketers.

## Measurement

The cricketdata package extracts match information from official sources, including the ESPN and Cricsheet databases, scorecards, and records maintained by cricket boards. These sources provide detailed records of each international match played by Pakistan and other cricketing teams worldwide. These matches represent the real-world phenomena we aim to analyze.

This real-world data is transformed into structured entries in our dataset through the manual recording of match information by officials during the game. Key details, such as match ID, participating teams, venue, date, event, toss winner/decision, player of the match, umpires, match winner, and more, are carefully documented. Once recorded, this information is made available for download on official websites. The datasets are then retrieved and imported into R using the cricketdata package.

This measurement approach is highly reliable, as it relies on official cricketing records. However, it also highlights the limitations inherent in working with aggregated datasets. For example, factors that can influence match outcomes—such as team morale or crowd effects—remain unquantified.

## Variable Analysis
Figures @fig-match-stats-1 and @fig-match-stats-2 display the count of several important match statistics. From these, we can observe that Pakistan faces a variety of opponents, with the most frequent ones being England, Sri Lanka, and Australia. Notably, Pakistan has had a challenging record against Australia, losing a significant number of matches, while many of their matches ended in draws.

Regarding coin toss outcomes, Sri Lanka and South Africa appear to have won a substantial number of tosses. However, it's important to note that while the graphs for match outcomes and tosses won may give the impression that Pakistan was victorious in most of these events, we must keep in mind that the data has been filtered to include only Pakistan's matches. This means Pakistan appears more frequently in the data set compared to other teams, which accounts for the higher statistics shown. These figures represent Pakistan's performance specifically, not that of other teams.

The following two graphs show the number of matches and tosses that Pakistan won or lost. The win-loss ratio is approximately 0.5, indicating that Pakistan has lost about twice as many matches as it has won. Additionally, the coin toss win ratio is close to even, reflecting a near 50/50 chance of Pakistan winning the toss.

Finally, the last graph illustrates the number of matches Pakistan played each year. We can see a peak in 2016, with a gradual decline in matches played over the subsequent years, culminating in a sharp drop in the most recent year—comparable to the period during the COVID-19 pandemic.

```{r}
#| label: fig-match-stats-1
#| fig-cap: Counts for match variables of interst
#| echo: false

data <- pakistan_match_data %>%
  mutate(pakistan_win = case_when(
    winner == "Pakistan" ~ 1,
    TRUE ~ 0))%>%
  mutate(toss_win = case_when(
    toss_winner == "Pakistan" ~ 1,
    TRUE ~ 0))%>%
  mutate(year = lubridate::year(date))

# Convert relevant columns to factors if they represent categorical data
data <- data %>%
  mutate(
    team1 = as.factor(team1),
    team2 = as.factor(team2),
    winner = as.factor(winner),
    toss_winner = as.factor(toss_winner),
    pakistan_win = as.factor(pakistan_win),
    toss_win = as.factor(toss_win)
  )

# Figure 2: Distribution of matches played by Team2
fig2 <- ggplot(data, aes(x = team2)) +
  geom_bar(fill = "coral") +
  labs(title = "Pakistan Opponent Team", x = "Team", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Figure 3: Match outcomes (Winner)
fig3 <- ggplot(data, aes(x = winner)) +
  geom_bar(fill = "darkgreen") +
  labs(title = "Matches Won by Each Team", x = "Team", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Figure 4: Distribution of Toss Winners
fig4 <- ggplot(data, aes(x = toss_winner)) +
  geom_bar(fill = "purple") +
  labs(title = "Tosses Won by each Team", x = "Team", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Figure 5: Pakistan Win Outcomes
fig5 <- ggplot(data, aes(x = pakistan_win)) +
  geom_bar(fill = "gold") +
  labs(title = "Matches Won by Pakistan", x = "Outcome", y = "Count") +
  scale_x_discrete(labels = c("0" = "Lost", "1" = "Won")) +  # Change 1 and 0 to Won and Lost
  theme_minimal()

# Figure 6: Toss Win Outcomes
fig6 <- ggplot(data, aes(x = toss_win)) +
  geom_bar(fill = "dodgerblue") +
  labs(title = "Tosses Won by Pakistan", x = "Outcome", y = "Count") +
  scale_x_discrete(labels = c("0" = "Lost", "1" = "Won")) +  # Change 1 and 0 to Won and Lost
  theme_minimal()

# Figure 7: Match Results by Year
fig7 <- ggplot(data, aes(x = year)) +
  geom_bar(fill = "brown") +
  labs(title = "Pakistan's Matches Played \nper Year", x = "Year", y = "Number of Matches") +
  theme_minimal()

combines_plot1 <- (fig2)/(fig3)
print(combines_plot1)

```

```{r}
#| label: fig-match-stats-2
#| fig-cap: Counts for match variables of interst
#| echo: false

combines_plot2 <- (fig4 | fig5) /( fig6 | fig7)
print(combines_plot2)
```

# Model {#sec-model}

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results {#sec-results}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# library(rstanarm)
# 
# first_model <-
#   readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false
# 
# modelsummary::modelsummary(
#   list(
#     "First model" = first_model
#   ),
#   statistic = "mad",
#   fmt = 2
# )
```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
# 
# pp_check(first_model) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(first_model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

# plot(first_model, "trace")
# 
# plot(first_model, "rhat")
```

\newpage

# References

### 2024-12-02 scripts/03-clean_data.R

@@ -44,7 +44,22 @@ cleaned_match_data <-
          winner_runs,
          toss_winner) |>
   mutate(date = as.character(date)) |>
-  mutate(date = ymd(date))  # Convert date to lubridate Date format
+  mutate(date = ymd(date)) |>
+  mutate(winner = ifelse(is.na(winner), "Draw", winner))  # Replace NA with "Draw" in winner column
+
+for (i in 1:nrow(cleaned_match_data)) {
+  if (cleaned_match_data$team1[i] == "Pakistan" || cleaned_match_data$team2[i] == "Pakistan") {
+    if(cleaned_match_data$team1[i] != "Pakistan"){
+      temp <- cleaned_match_data$team1[i]
+      cleaned_match_data$team1[i] <- "Pakistan"
+      cleaned_match_data$team2[i] <- temp
+    }
+  }
+}
+
+# Filter match data for Pakistan
+pakistan_matches <- cleaned_match_data %>%
+  filter(team1 == "Pakistan")
 
 cleaned_player_data <- bind_rows(
   raw_bat_data %>% mutate(dataset_type = "bat"),
@@ -106,5 +121,7 @@ cleaned_player_data <- bind_rows(
 #### Save data ####
 write_parquet(cleaned_match_data,
               "data/02-analysis_data/cleaned_match_data.parquet")
+write_parquet(pakistan_matches,
+              "data/02-analysis_data/cleaned_pakistan_match_data.parquet")
 write_parquet(cleaned_player_data,
               "data/02-analysis_data/cleaned_player_data.parquet")

### 2024-12-02 scripts/06-model_data.R

@@ -41,7 +41,6 @@ pakistan_matches <- match_data %>%
   mutate(toss_win = case_when(
     toss_winner == "Pakistan" ~ 1,
     TRUE ~ 0))%>%
-  select(-toss_decision)%>%
   mutate(year = lubridate::year(date))
 
 

### 2024-12-04 paper/paper.qmd
Added results section to paper

@@
 toc-depth: 2
+editor: 
+  markdown: 
+    wrap: 72
 ---
@@
 library(arrow)
-library(patchwork)
+library(modelsummary)
+library(tinytable)
+library(knitr)
+library(kableExtra)
+
+player_data <- read_parquet("data/02-analysis_data/cleaned_player_data.parquet")
 
 pakistan_match_data <- read_parquet(here("data/02-analysis_data/cleaned_pakistan_match_data.parquet"))
+model <- readRDS(here("models/win_chance.rds"))
+confusion_matrix <- readRDS(here("models/confusion_matrix_win_chance.rds"))
+
 ```
@@
 # Introduction
-Pakistan’s cricket team holds a storied legacy, marked by remarkable highs and challenging lows.
-From their 1992 World Cup triumph to fluctuations in form over the decades, analyzing trends in performance offers insights into their journey and areas for strategic improvement.
 
-Despite an extensive focus on player statistics and match highlights, there remains a lack of systematic statistical analysis to evaluate long-term performance trends.
-This paper fills this gap by analyzing data from 2002 to 2024 to understand what factors contribute to Pakistan’s success in international cricket.
+Pakistan’s cricket team holds a storied legacy, marked by remarkable
+highs and challenging lows. From their 1992 World Cup triumph to
+fluctuations in form over the decades, analyzing trends in performance
+offers insights into their journey and areas for strategic improvement.
 
-We employ a linear regression model to analyze match outcomes, considering variables like year, opponent, and toss outcomes.
-The focus is on identifying trends and predicting the probability of victory under varying conditions.
-Our findings not only contribute to the literature on sports analytics but also provide actionable insights for coaches, analysts, and cricket administrators not only in Pakistan, but around the world.
+Despite an extensive focus on player statistics and match highlights,
+there remains a lack of systematic statistical analysis to evaluate
+long-term performance trends. This paper fills this gap by analyzing
+data from 2002 to 2024 to understand what factors contribute to
+Pakistan’s success in international cricket.
 
-The structure of this paper is as follows: @sec-data discusses the data types included in the raw data, the cleaning process for the data, and the reason for selecting the data set we did.
-@sec-model discusses model specification and justification for a Linear Regression model.
-@sec-results analyzes the trends and correlations between different variables utilizing tabular and graphical means.
-@sec-discussion discusses the results of @sec-results going into detail on what the simulation results can tell us about Pakistan cricket performance, as well as assumptions and limitations in data.
+We employ a linear regression model to analyze match outcomes,
+considering variables like year, opponent, and toss outcomes. The focus
+is on identifying trends and predicting the probability of victory under
+varying conditions. Our findings not only contribute to the literature
+on sports analytics but also provide actionable insights for coaches,
+analysts, and cricket administrators not only in Pakistan, but around
+the world.
+
+The structure of this paper is as follows: @sec-data discusses the data
+types included in the raw data, the cleaning process for the data, and
+the reason for selecting the data set we did. @sec-model discusses model
+specification and justification for a Linear Regression model.
+@sec-results analyzes the trends and correlations between different
+variables utilizing tabular and graphical means. @sec-discussion
+discusses the results of @sec-results going into detail on what the
+simulation results can tell us about Pakistan cricket performance, as
+well as assumptions and limitations in data.
 
 ## Estimand
-The estimand is the probability that Pakistan wins a cricket match, given the opponent, toss outcome, and year.
+
+The estimand is the probability that Pakistan wins a cricket match,
+given the opponent, toss outcome, and year.
 
@@
 
-The data for this study were obtained from the cricketdata R package [@cricketdata] and all analysis was performed using `R` [@citeR] alongside the following packages: `tibble` [@tibble], `readr` [@readr], `arrow` [@arrow], `tidyverse` [@tidyverse], `dplyr` [@dplyr], `testthat` [@testthat], `caret` [@caret], `rstanarm` [@rstanarm], `tidyr` [@tidyr], patchwork [@patchwork], and ggplot2 [@ggplot2].
+The data for this study were obtained from the cricketdata R package
+[@cricketdata] and all analysis was performed using `R` [@citeR]
+alongside the following packages: `tibble` [@tibble], `readr` [@readr],
+`arrow` [@arrow], `tidyverse` [@tidyverse], `dplyr` [@dplyr], `testthat`
+[@testthat], `caret` [@caret], `rstanarm` [@rstanarm], `tidyr` [@tidyr],
+modelsummary [@modelsummary], tinytable [@tinytable], knitr [@knitr], and ggplot2 [@ggplot2].
 
-As mentioned before the raw data was collected from the cricketdata library which contains data on international and other major cricket matches sourced from ESPN Cricinfo [@espncricinfo] and Cricsheet [@cricsheet].
-It contains details on the individual match year, teams, result, and toss winner.
-Additionally, the data set features comprehensive career statistics for individual players, including total runs, total balls faced, total wickets, and more.
-This data set was chosen for its ease of implementation within R, as it is available as a package, and for its ability to scrape data from multiple reputable cricket sources, such as ESPN, thereby enhancing the overall reliability of the data set.
+As mentioned before the raw data was collected from the cricketdata
+library which contains data on international and other major cricket
+matches sourced from ESPN Cricinfo [@espncricinfo] and Cricsheet
+[@cricsheet]. It contains details on the individual match year, teams,
+result, and toss winner. Additionally, the data set features
+comprehensive career statistics for individual players, including total
+runs, total balls faced, total wickets, and more. This data set was
+chosen for its ease of implementation within R, as it is available as a
+package, and for its ability to scrape data from multiple reputable
+cricket sources, such as ESPN, thereby enhancing the overall reliability
+of the data set.
 
 ## Raw Data
-The cricketdata library offers a variety of functions that provide data sets on different cricket statistics.
-However, for our analysis, we focus on two key areas: individual match statistics, to determine which matches Pakistan played in and how they performed, and player career statistics, to assess how the quality of players has evolved over time and its impact on Pakistan's overall performance.
-To extract the relevant data, we primarily used two functions from the library: fetch_cricinfo, which retrieves individual career performance data, and fetch_cricsheet, which provides detailed match information.
+
+The cricketdata library offers a variety of functions that provide data
+sets on different cricket statistics. However, for our analysis, we
+focus on two key areas: individual match statistics, to determine which
+matches Pakistan played in and how they performed, and player career
+statistics, to assess how the quality of players has evolved over time
+and its impact on Pakistan's overall performance. To extract the
+relevant data, we primarily used two functions from the library:
+fetch_cricinfo, which retrieves individual career performance data, and
+fetch_cricsheet, which provides detailed match information.
 
 ## Data Cleaning
-The raw match data initially contained 820 observations across 25 variables, each detailing a match between two countries.
-The data cleaning process involved selecting only matches in which Pakistan played, and retaining only the variables relevant to our analysis.
-These variables include:
+
+The raw match data initially contained 820 observations across 25
+variables, each detailing a match between two countries. The data
+cleaning process involved selecting only matches in which Pakistan
+played, and retaining only the variables relevant to our analysis. These
+variables include:
 
@@
 -   `Date` - The date of the match (eg. 2008-01-02)
-- `Winner` - The team that won the match between Team1 and Team2 (NA if the match was a draw).
-- `Winner_wickets` - The number of wickets the winning team took (NA if the match was a draw).
-- `Winner_runs` - The number of runs the winning team scored (NA if the match was a draw).
+-   `Winner` - The team that won the match between Team1 and Team2 (NA
+    if the match was a draw).
+-   `Winner_wickets` - The number of wickets the winning team took (NA
+    if the match was a draw).
+-   `Winner_runs` - The number of runs the winning team scored (NA if
+    the match was a draw).
 -   `Toss_winner` - The team that won the coin toss.
-- `Toss_decision` - The decision made by the toss winner, indicating whether they chose to field or bowl.
+-   `Toss_decision` - The decision made by the toss winner, indicating
+    whether they chose to field or bowl.
 
-The raw player career data was broken into batting, bowling, and fielding categories with some players appearing in multiple data sets, thus the cleaning process involved combining the 3 data sets and ensuring proper handling of one player in each set.
-After this, only Pakistani cricket players were selected and variables which were important to use were retained.
-These variables are:
+The raw player career data was broken into batting, bowling, and
+fielding categories with some players appearing in multiple data sets,
+thus the cleaning process involved combining the 3 data sets and
+ensuring proper handling of one player in each set. After this, only
+Pakistani cricket players were selected and variables which were
+important to use were retained. These variables are:
 
@@
 -   `End` - The year in which the player's career concluded.
-- `Matches` - The total number of matches the player participated in throughout their career.
-- `Bat_innings` -  The total number of batting innings the player participated in.
-- `Bowl_innings` -  The total number of bowling innings the player participated in.
-- `Field_innings` -  The total number of fielding innings the player participated in.
-- `Bat_runs` - The total number of runs the player scored while batting.
-- `Not_puts` - The number of times the player remained not out while batting.
-- `Bat_average` - The player's batting average, calculated as the total number of runs scored divided by the number of times they were dismissed.
-- `Bowl-runs` - The total number of runs conceded by the player while bowling. This statistic reflects how many runs the player allowed the opposition to score during their bowling spells.
-- `Wickets` - The total number of wickets taken by the player in their career.
-- `Economy` - The average number of runs the player concedes per over bowled.
-It is calculated by dividing the total number of runs given by the total number of overs bowled.
-A lower economy rate typically indicates a bowler who is more effective at restricting the opposition's scoring.
-- `Dismissals` - The total number of dismissals the player has contributed to as a fielder or wicketkeeper.
+-   `Matches` - The total number of matches the player participated in
+    throughout their career.
+-   `Bat_innings` - The total number of batting innings the player
+    participated in.
+-   `Bowl_innings` - The total number of bowling innings the player
+    participated in.
+-   `Field_innings` - The total number of fielding innings the player
+    participated in.
+-   `Bat_runs` - The total number of runs the player scored while
+    batting.
+-   `Not_puts` - The number of times the player remained not out while
+    batting.
+-   `Bat_average` - The player's batting average, calculated as the
+    total number of runs scored divided by the number of times they were
+    dismissed.
+-   `Bowl-runs` - The total number of runs conceded by the player while
+    bowling. This statistic reflects how many runs the player allowed
+    the opposition to score during their bowling spells.
+-   `Wickets` - The total number of wickets taken by the player in their
+    career.
+-   `Economy` - The average number of runs the player concedes per over
+    bowled. It is calculated by dividing the total number of runs given
+    by the total number of overs bowled. A lower economy rate typically
+    indicates a bowler who is more effective at restricting the
+    opposition's scoring.
+-   `Dismissals` - The total number of dismissals the player has
+    contributed to as a fielder or wicketkeeper.
 
-After applying these cleaning procedures, we were left with 151 observations across 7 variables for the matches Pakistan has played in, and 344 observations across 14 variables for the career information of individual Pakistani cricketers.
+After applying these cleaning procedures, we were left with 151
+observations across 7 variables for the matches Pakistan has played in,
+and 344 observations across 14 variables for the career information of
+individual Pakistani cricketers.
 
@@
 
-The cricketdata package extracts match information from official sources, including the ESPN and Cricsheet databases, scorecards, and records maintained by cricket boards.
-These sources provide detailed records of each international match played by Pakistan and other cricketing teams worldwide.
-These matches represent the real-world phenomena we aim to analyze.
+The cricketdata package extracts match information from official
+sources, including the ESPN and Cricsheet databases, scorecards, and
+records maintained by cricket boards. These sources provide detailed
+records of each international match played by Pakistan and other
+cricketing teams worldwide. These matches represent the real-world
+phenomena we aim to analyze.
 
-This real-world data is transformed into structured entries in our dataset through the manual recording of match information by officials during the game.
-Key details, such as match ID, participating teams, venue, date, event, toss winner/decision, player of the match, umpires, match winner, and more, are carefully documented.
-Once recorded, this information is made available for download on official websites.
-The datasets are then retrieved and imported into R using the cricketdata package.
+This real-world data is transformed into structured entries in our
+dataset through the manual recording of match information by officials
+during the game. Key details, such as match ID, participating teams,
+venue, date, event, toss winner/decision, player of the match, umpires,
+match winner, and more, are carefully documented. Once recorded, this
+information is made available for download on official websites. The
+datasets are then retrieved and imported into R using the cricketdata
+package.
 
-This measurement approach is highly reliable, as it relies on official cricketing records.
-However, it also highlights the limitations inherent in working with aggregated datasets.
-For example, factors that can influence match outcomes—such as team morale or crowd effects—remain unquantified.
+This measurement approach is highly reliable, as it relies on official
+cricketing records. However, it also highlights the limitations inherent
+in working with aggregated datasets. For example, factors that can
+influence match outcomes—such as team morale or crowd effects—remain
+unquantified.
 
 ## Variable Analysis
-Figures @fig-match-stats-1 and @fig-match-stats-2 display the count of several important match statistics.
-From these, we can observe that Pakistan faces a variety of opponents, with the most frequent ones being England, Sri Lanka, and Australia.
-Notably, Pakistan has had a challenging record against Australia, losing a significant number of matches, while many of their matches ended in draws.
 
-Regarding coin toss outcomes, Sri Lanka and South Africa appear to have won a substantial number of tosses.
-However, it's important to note that while the graphs for match outcomes and tosses won may give the impression that Pakistan was victorious in most of these events, we must keep in mind that the data has been filtered to include only Pakistan's matches.
-This means Pakistan appears more frequently in the data set compared to other teams, which accounts for the higher statistics shown.
-These figures represent Pakistan's performance specifically, not that of other teams.
+Figures @fig-match-stats displays the count of several important match
+statistics. From these, we can observe that Pakistan faces a variety of
+opponents, with the most frequent ones being England, Sri Lanka, and
+Australia. Notably, Pakistan has had a challenging record against
+Australia, losing a significant number of matches, while many of their
+matches ended in draws.
 
-The following two graphs show the number of matches and tosses that Pakistan won or lost.
-The win-loss ratio is approximately 0.5, indicating that Pakistan has lost about twice as many matches as it has won.
-Additionally, the coin toss win ratio is close to even, reflecting a near 50/50 chance of Pakistan winning the toss.
+Regarding coin toss outcomes, Sri Lanka and South Africa appear to have
+won a substantial number of tosses. However, it's important to note that
+while the graphs for match outcomes and tosses won may give the
+impression that Pakistan was victorious in most of these events, we must
+keep in mind that the data has been filtered to include only Pakistan's
+matches. This means Pakistan appears more frequently in the data set
+compared to other teams, which accounts for the higher statistics shown.
+These figures represent Pakistan's performance specifically, not that of
+other teams.
 
-Finally, the last graph illustrates the number of matches Pakistan played each year.
-We can see a peak in 2016, with a gradual decline in matches played over the subsequent years, culminating in a sharp drop in the most recent year—comparable to the period during the COVID-19 pandemic.
+The following two graphs show the number of matches and tosses that
+Pakistan won or lost. The win-loss ratio is approximately 0.5,
+indicating that Pakistan has lost about twice as many matches as it has
+won. Additionally, the coin toss win ratio is close to even, reflecting
+a near 50/50 chance of Pakistan winning the toss.
+
+Finally, the last graph illustrates the number of matches Pakistan
+played each year. We can see a peak in 2016, with a gradual decline in
+matches played over the subsequent years, culminating in a sharp drop in
+the most recent year—comparable to the period during the COVID-19
+pandemic.
 
 ```{r}
-#| label: fig-match-stats-1
+#| label: fig-match-stats
 #| fig-cap: Counts for match variables of interst
+#| fig-subcap: ["Pakistan Opponent Team","Matches Won by Each Team",
+#| "Tosses Won by each Team", "Matches Won by Pakistan", "Tosses Won by Pakistan", "Pakistan's Matches Played per Year"]
+#| layout-ncol: 2
+#| layout-nrow: 3
 #| echo: false
@@
 
-# Figure 2: Distribution of matches played by Team2
-fig2 <- ggplot(data, aes(x = team2)) +
+# Figure 1: Distribution of matches played by Team2
+ggplot(data, aes(x = team2)) +
   geom_bar(fill = "coral") +
-  labs(title = "Pakistan Opponent Team", x = "Team", y = "Count") +
+  labs(x = "Team", y = "Count") +
   theme_minimal() +
@@
 
-# Figure 3: Match outcomes (Winner)
-fig3 <- ggplot(data, aes(x = winner)) +
+# Figure 2: Match outcomes (Winner)
+ggplot(data, aes(x = winner)) +
   geom_bar(fill = "darkgreen") +
-  labs(title = "Matches Won by Each Team", x = "Team", y = "Count") +
+  labs(x = "Team", y = "Count") +
   theme_minimal() +
@@
 
-# Figure 4: Distribution of Toss Winners
-fig4 <- ggplot(data, aes(x = toss_winner)) +
+# Figure 3: Distribution of Toss Winners
+ggplot(data, aes(x = toss_winner)) +
   geom_bar(fill = "purple") +
-  labs(title = "Tosses Won by each Team", x = "Team", y = "Count") +
+  labs(x = "Team", y = "Count") +
   theme_minimal() +
@@
 
-# Figure 5: Pakistan Win Outcomes
-fig5 <- ggplot(data, aes(x = pakistan_win)) +
+# Figure 4: Pakistan Win Outcomes
+ggplot(data, aes(x = pakistan_win)) +
   geom_bar(fill = "gold") +
-  labs(title = "Matches Won by Pakistan", x = "Outcome", y = "Count") +
+  labs(x = "Outcome", y = "Count") +
   scale_x_discrete(labels = c("0" = "Lost", "1" = "Won")) +  # Change 1 and 0 to Won and Lost
@@
 
-# Figure 6: Toss Win Outcomes
-fig6 <- ggplot(data, aes(x = toss_win)) +
+# Figure 5: Toss Win Outcomes
+ggplot(data, aes(x = toss_win)) +
   geom_bar(fill = "dodgerblue") +
-  labs(title = "Tosses Won by Pakistan", x = "Outcome", y = "Count") +
+  labs(x = "Outcome", y = "Count") +
   scale_x_discrete(labels = c("0" = "Lost", "1" = "Won")) +  # Change 1 and 0 to Won and Lost
@@
 
-# Figure 7: Match Results by Year
-fig7 <- ggplot(data, aes(x = year)) +
+# Figure 6: Match Results by Year
+ggplot(data, aes(x = year)) +
   geom_bar(fill = "brown") +
-  labs(title = "Pakistan's Matches Played \nper Year", x = "Year", y = "Number of Matches") +
+  labs(x = "Year", y = "Number of Matches") +
   theme_minimal()
 
-combines_plot1 <- (fig2)/(fig3)
-print(combines_plot1)
-
-```
-
-```{r}
-#| label: fig-match-stats-2
-#| fig-cap: Counts for match variables of interst
-#| echo: false
-
-combines_plot2 <- (fig4 | fig5) /( fig6 | fig7)
-print(combines_plot2)
 ```
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+I used a Bayesian Logistic Regression model to estimate the probability
+of Pakistan winning a cricket match based on several key factors.
+Logistic regression is a statistical technique used for binary
+classification, where the goal is to predict the probability of a
+categorical outcome. In this case, the outcome is whether Pakistan wins
+or loses a match.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+## Model Overview
 
-## Model set-up
+The model incorporates three independent variables that may influence
+match outcomes: year, opponent, and toss winner. The dependent variable
+is match_outcome, where 1 indicates a win for Pakistan and 0 indicates a
+loss. This setup allows us to examine the influence of key variables on
+Pakistan's performance.
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+By incorporating these variables, the model explores how factors such as
+the year of the match, the opposing team, and whether Pakistan won the
+toss (which influences the decision to bat or bowl) affect Pakistan’s
+likelihood of winning. This approach helps uncover trends in Pakistan's
+performance over time and under varying conditions, providing insights
+into factors that may inform future match strategies.
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+## Model Equation
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+The model I will be using can be described using the following equation:
 
-### Model justification
+```{=tex}
+\begin{equation}
+\log\left(\frac{\hat{p}}{1 - \hat{p}}\right) = \beta_0 + \beta_1 \times \text{year} + \beta_2 \times \text{opponent} + \beta_3 \times \text{toss winner} 
+\end{equation}
+```
+$$
+\begin{aligned}
+\beta_0 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_1 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_2 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_3 & \sim \mbox{Normal}(0, 2.5)
+\end{aligned}
+$$ where,
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+-   $\hat{p}$ is the predicted probability of Pakistan winning a match.
+-   $\beta_0$ is the intercept (the log-odds of winning when all
+    predictors are at their reference level).
+-   $\beta_1$ is the coefficient for the year variable, representing how
+    match outcomes are affected by the year of the match.
+-   $\beta_2$ are the coefficients for each level of the categorical
+    variable team2 (the opponent), with the reference category excluded.
+    This allows the model to compare the odds of winning against
+    different opponents.
+-   $\beta_3$ is the coefficient for the toss_win variable (whether
+    Pakistan won the toss).
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+## Model Priors
+
+The priors used in the model represent our assumptions about the
+parameters before observing the data. These priors are chosen based on
+reasonable beliefs about the likely values of the coefficients and are
+specified as follows:
+
+-   `Prior for the Coefficients (year, team2, toss_win)`: Each of the
+    model coefficients is assigned a Normal prior with a mean of 0 and a
+    standard deviation of 2.5. This reflects the assumption that we
+    expect the effects of the predictors to be centered around zero,
+    with a reasonable amount of uncertainty (a wide spread due to the
+    scale of 2.5) about their true values.
+
+-   `Prior for the Intercept`: A normal prior with the same parameters
+    as the coefficients (mean = 0, scale = 2.5) is used for the
+    intercept. The intercept represents the baseline log-odds of
+    Pakistan winning a match when all other predictors are zero.
+
+-   `Prior for the Auxiliary Parameter (prior_aux)`: This parameter
+    represents the dispersion of the logistic regression model. An
+    Exponential prior with a rate of 1 is used, which is typical in
+    Bayesian models for logistic regression. The Exponential prior
+    allows for a distribution with only non-negative values, reflecting
+    uncertainty in the variance of match outcomes, with higher values
+    suggesting more variability in outcomes.
+
+The Bayesian approach incorporates prior knowledge (through these
+priors) and updates it with the data from the training set, enabling the
+model to estimate relationships while accounting for uncertainty in the
+coefficients.
+
+## Model Justification
+
+A decision tree model was also considered, which could capture
+non-linear relationships and interactions between predictors. However,
+decision trees are more prone to overfitting and can be less
+interpretable than logistic regression, especially in cases with
+categorical variables like team2.
+
+An alternative approach considered was a frequentist logistic regression
+model. This model would not incorporate prior information and would
+instead focus purely on the likelihood of the data. However, the
+Bayesian approach was chosen due to its ability to incorporate prior
+knowledge and handle uncertainty more effectively, especially in the
+context of small sample sizes and the complexity of cricket data.
+
+This model provides a useful framework for estimating Pakistan's
+likelihood of winning a match based on key factors. By using a Bayesian
+Logistic Regression approach, the model incorporates uncertainty in the
+estimates, offers a clear interpretation of the effects of each
+predictor, and is flexible enough to adapt to new data as it becomes
+available. However, it should be acknowledged that there are limitations
+regarding omitted variables and potential model assumptions that could
+be addressed in future iterations.
+
+This is an overview of the model. More details such as assumptions,
+limitations, convergence, and validation can be found in
+@sec-model-details
 
@@
 
-Our results are summarized in @tbl-modelresults.
+Our results are summarized in @tbl-model-results, and @fig-model-results
+presents the predicted values of the coefficients. This figure also
+includes the 95% confidence intervals for the predictors, as estimated
+by our Bayesian model. A 95% confidence interval indicates that there is
+a 95% probability that the true value of the parameter lies within the
+specified range, given the observed data. In the graph, each point
+represents the predicted value, and the horizontal line extending from
+each point shows the 95% confidence interval.
+
+Variables to the right of 0 have a positive correlation with the
+outcome, meaning that increasing these values increases the probability
+of Pakistan winning. Conversely, variables to the left of 0 exhibit a
+negative correlation, meaning that increasing these variables reduces
+the probability of Pakistan winning a match.
+
+Finally, it is important to note that @fig-model-results is plotted
+without the intercept term and the team2Ireland term. As shown in
+@tbl-model-results, these two values have much larger magnitudes
+compared to the other predictors. If included in the plot, their scale
+would be so large that the smaller variations in the other coefficients
+would be obscured.
 
 ```{r}
+#| label: tbl-model-results
+#| tbl-cap: "Coefficient estimates and standard errors"
 #| echo: false
@@
 #| warning: false
-#| message: false
 
-# library(rstanarm)
-# 
-# first_model <-
-#   readRDS(file = here::here("models/first_model.rds"))
+ modelsummary::modelsummary(
+   list("Model" = model),
+   statistic = "mad",
+   fmt = 2,
+   gof_omit = "Num.Obs.|R2|Log.Lik.|RMSE|ELPD|LOOIC|LOOIC s.e.|WAIC"
+ )
 ```
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| label: fig-model-results
+#| fig-cap: "Coefficient estimates and 95% confidence intervals"
 #| warning: false
-# 
-# modelsummary::modelsummary(
-#   list(
-#     "First model" = first_model
-#   ),
-#   statistic = "mad",
-#   fmt = 2
-# )
+
+ modelplot(model, coef_omit = "Intercept|team2Ireland") + 
+   labs(x = "Coefficient Estimate", y="Predictor Variable")
 ```
+
+Analyzing some of these coefficients further, The results from the
+regression model are as follows:
+
+-   **Intercept**: The intercept of -55.94 (with a standard error of
+    80.63) represents the log-odds of Pakistan winning a match when all
+    other variables are set to their reference categories. This value is
+    quite large in magnitude, but it is not directly interpretable in
+    isolation. The negative value suggests a very low baseline
+    probability of winning in the absence of other influencing factors.
+
+-   **Year**: The coefficient for year is 0.03 (standard error = 0.04).
+    This positive coefficient suggests that, as time progresses, the
+    probability of Pakistan winning a match increases very slightly.
+    However, the small magnitude of the coefficient indicates that this
+    effect is minimal. The relatively narrow standard error suggests
+    that the estimate is reasonably precise, but the practical
+    significance of this trend is modest.
+
+-   **Opponent**: The coefficients for the opponent variable represent
+    how the likelihood of Pakistan winning a match changes when facing
+    different teams. These coefficients should be interpreted relative
+    to the reference team, which is not explicitly listed here but is
+    assumed to be the team with the lowest or most neutral effect:
+
+    -   `team2Bangladesh` has a coefficient of 3.62 (standard error =
+        1.31), which indicates a significant positive association with
+        Pakistan’s success. Matches against Bangladesh appear to
+        increase Pakistan's chances of winning.
+
+    -   `team2Sri Lanka` shows a positive coefficient of 1.09 (standard
+        error = 0.72), suggesting that matches against Sri Lanka are
+        associated with a higher probability of Pakistan winning
+        compared to the reference team, though the effect is less
+        pronounced than for Bangladesh.
+
+    -   The coefficients for other teams, such as `team2India` (0.24, standard error = 1.44) and `team2South Africa` (0.29, standard error = 0.90), are much smaller, indicating weaker associations with match success.
+    - `team2Ireland` has the largest coefficient (19.77, standard error = 15.67), indicating a very high probability of success when playing against Ireland.
+However, this coefficient also has a large standard error, meaning there is significant uncertainty around this estimate.
+
+- `Toss Win`: The coefficient for toss_win is 0.60 (standard error = 0.45), suggesting that winning the toss slightly increases Pakistan's likelihood of winning the match.
+This result supports the common understanding that the team winning the toss often gains an advantage in deciding whether to bat or bowl first.
+
+```{r}
+#| label: fig-opponent-stats
+#| fig-cap: "Pakistan's wins and losses against all teams"
+#| fig-subcap: ["Sri Lanka","New Zealand", "Australia", "England", "South Africa", "West Indies", "Zimbabwe", "Bangladesh", "Ireland", "India"]
+#| layout-ncol: 2
+#| layout-nrow: 5
+#| echo: false
+#| eval: true
+#| warning: false
+
+# Create a new column that accounts for wins, losses, and draws
+pakistan_match_data <- pakistan_match_data %>%
+  mutate(pakistan_result = case_when(
+    winner == "Pakistan" ~ "Win",
+    winner == "Draw" ~ "Draw",
+    TRUE ~ "Loss"
+  ))
+
+# Loop over each opponent team
+opponents <- unique(pakistan_match_data$team2)  # Get all unique opponents
+
+# Create and save a plot for each opponent
+for (opponent in opponents) {
+
+  # Filter the data for the current opponent
+  filtered_matches <- pakistan_match_data %>% filter(team2 == opponent)
+
+  # Summarize the number of wins, losses, and draws
+  win_loss_summary <- filtered_matches %>%
+    group_by(pakistan_result) %>%
+    summarise(count = n(), .groups = "drop")
+  ggplot(data, aes(x = team2)) +
+  geom_bar(fill = "coral") +
+  labs(x = "Team", y = "Count") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  # Create the plot
+  plot <- ggplot(win_loss_summary, aes(x = pakistan_result, y = count, fill = pakistan_result)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(
+      x = "Result",
+      y = "Count",
+      fill = "Result"
+    ) +
+    scale_fill_manual(values = c("Win" = "darkgreen", "Loss" = "brown2", "Draw" = "darkgoldenrod2")) +  # Custom colors for Win, Loss, Draw
+    coord_cartesian(ylim = c(0, 14)) +
+    theme_minimal()+
+    theme(legend.position = "none")
+
+print(plot)
+}
+```
+@fig-opponent-stats illustrates the distribution of match outcomes (wins, losses, and draws) for Pakistan against eight different cricket teams: Sri Lanka, New Zealand, England, Australia, South Africa, West Indies, Zimbabwe, and Bangladesh.
+These results provide an overview of Pakistan's historical performance against each team and highlight patterns in their match outcomes.
+
+- `Sri Lanka` (Figure 3a): Pakistan's matches against Sri Lanka are fairly balanced, with a slightly higher frequency of losses compared to wins.
+Draws are the least frequent outcome, suggesting a competitive dynamic between the two teams.
+
+- `New Zealand` (Figure 3b): Matches against New Zealand reveal a similar trend, with losses being more frequent than wins.
+The proportion of draws remains relatively low, indicating fewer instances of matches being evenly contested.
+
+- `England` (Figure 3c): Pakistan has a notably higher number of losses against England, with draws constituting the smallest category.
+This indicates that England has historically been a challenging opponent for Pakistan.
+
+- `Australia` (Figure 3d): A significant disparity is observed in matches against Australia, where losses outnumber wins by a considerable margin.
+This trend underscores the dominance of Australia in these encounters.
+
+- `South Africa` (Figure 3e): Matches against South Africa also show a higher proportion of losses compared to wins.
+However, the gap between these two outcomes is less pronounced compared to matches against Australia.
+
+- `West Indies` (Figure 3f): Pakistan exhibits a stronger performance against West Indies, with wins slightly outnumbering losses.
+Draws remain the least frequent outcome, emphasizing decisive results in these encounters.
+
+- `Zimbabwe` (Figure 3g): Against Zimbabwe, Pakistan demonstrates a clear dominance, with wins significantly outnumbering losses.
+Draws are nearly negligible, highlighting Pakistan's consistent success in these matches.
+
+- `Bangladesh` (Figure 3h): Similar to Zimbabwe, Pakistan has a markedly higher number of wins against Bangladesh, with losses and draws being infrequent.
+This dominance underscores Pakistan's strong track record against this opponent.
+
+Overall, these graphs provide valuable insights into Pakistan's historical performance trends across different opponents.
+While some teams, such as Australia and England, have posed significant challenges, others, like Bangladesh and Zimbabwe, have been more favorable matchups for Pakistan.
+These findings align with the regression model coefficients presented earlier, where the opponent variable was a significant predictor of Pakistan's success in matches.
+For instance, the high win rates against Bangladesh and Zimbabwe correspond to their positive regression coefficients, while the challenging outcomes against teams like Australia reflect negative or smaller coefficients.
+
+```{r}
+#| label: fig-pk-wins
+#| fig-cap: "Number of Pakistan's Wins Over the Years"
+#| echo: false
+#| eval: true
+#| warning: false
+
+pakistan_win_data <- pakistan_match_data %>%
+    mutate(pakistan_win = case_when(
+    winner == "Pakistan" ~ 1,
+    TRUE ~ 0))%>%
+  mutate(year = lubridate::year(date))
+
+# Summarize the number of wins per year
+wins_by_year <- pakistan_win_data %>%
+  group_by(year) %>%
+  summarise(total_wins = sum(pakistan_win))
+
+# Create a bar plot showing the number of wins per year
+ggplot(wins_by_year, aes(x = year, y = total_wins)) +
+  geom_bar(stat = "identity", fill = "darkblue", show.legend = FALSE) +  # Set a fixed color for all bars
+  labs(
+    x = "Year",
+    y = "Total Wins"
+  ) +
+  theme_minimal()
+
+```
+@fig-pk-wins displays the number of wins Pakistan has achieved each year.
+From the graph, we observe that Pakistan had a modest performance from 2002 to 2010, with only 2-4 wins per year, followed by a sharp increase in 2011.
+After this surge, the team maintained an average of 4 wins per year, with some fluctuations.
+However, in 2019, there was a notable decline, with just one win recorded.
+Since this drop, Pakistan has struggled to reach their previous high of 4 wins per year, and the current average stands at approximately 2 wins per year.
 
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+If my paper were 10 pages, then should be be at least 2.5 pages. The
+discussion is a chance to show off what you know and what you learnt
+from all this.
 
@@
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+Please don't use these as sub-heading labels - change them to be what
+your point actually is.
 
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 
@@
 
-## Posterior predictive check
+## Model assumptions
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+-   `Independence of Observations`: We assume that the results of each
+    match are independent. This assumption may be violated in cases
+    where teams play multiple matches within a short period, but the
+    model does not account for such temporal dependencies.
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+-   `Logistic Relationship`: The model assumes a linear relationship
+    between the predictors (year, opponent, toss win) and the log-odds
+    of Pakistan winning. This assumption may not capture more complex
+    non-linear relationships, which would require more sophisticated
+    models, such as generalized additive models (GAMs).
+
+-   `No Interaction Effects`: The model currently does not include
+    interaction terms (e.g., how the effect of winning the toss might
+    differ by opponent). Interaction effects could be considered if they
+    were hypothesized to be important, but they are omitted here for
+    simplicity.
+
+-   `Reference Category for team2`: The model assumes that the effect of
+    playing against a reference opponent (the one excluded from the
+    dummy variables for team2) is the baseline against which the other
+    teams are compared. The choice of reference category is important
+    and could influence the interpretation of the coefficients.
+
+## Model limitations
+
+-   `Omitted Variables`: The model does not account for other
+    potentially important variables, such as player injuries, weather
+    conditions, or venue specifics. These factors could influence the
+    match outcome but are not included in the current model.
+
+-   `Data Representation`: The model is based on matches involving
+    Pakistan, so it does not generalize to other teams or international
+    cricket broadly. Additionally, the dataset may have biases (e.g., a
+    higher number of matches against certain opponents), which could
+    affect the accuracy of the predictions.
+
+-   `Data Sparsity`: The dataset available for analysis may not be large
+    enough to fully capture all the complexities of cricket matches,
+    especially with respect to less frequent events or rare outcomes.
+    This could lead to uncertainty in the model estimates, particularly
+    for smaller categories or outcomes that appear less frequently.
+
+-   `Data Granularity`: The model uses year as a continuous variable.
+    However, time trends in cricket might be better captured with more
+    granular temporal variables (e.g., by series or tournament), as the
+    year-by-year analysis may mask shorter-term trends.
+
+## Model validation
+
+We split the data into training and test sets to evaluate the predictive
+performance of the models. The first model was trained on a subset of
+the data, and its predictions were tested on the remaining matches in
+the test set. Specifically, we used out-of-sample testing, where the
+model was trained on the training set and evaluated on the test set to
+assess its ability to generalize to new, unseen data.
+
+For the first model, we predicted the probability of a match outcome
+(e.g., whether Pakistan would win) for each match in the test set. Using
+a threshold of 0.5, we classified these probabilities into binary
+outcomes (1 for a predicted win, 0 for a predicted loss). To evaluate
+model performance, we constructed a confusion matrix comparing the
+predicted classes to the actual outcomes of the test set. From this
+confusion matrix, we calculated accuracy as the proportion of correct
+predictions (i.e., the sum of true positives and true negatives divided
+by the total number of predictions). This provided an overall measure of
+how well the model performed in predicting match outcomes. The result of
+these tests is summarized in @tbl-model-metrics in the form of a
+confusion matrix and accuracy.
+
+We repeated the same procedure for the second model, predicting the
+match outcome based on a different set of features (e.g., year and
+opponent). The accuracy for both models was computed and reported
+separately.
+
+This approach provided a robust evaluation of each model's predictive
+performance on unseen data, with the confusion matrix and accuracy
+offering insights into the model’s classification capabilities. While
+accuracy provides a good indication of overall model performance, we
+also recognize that other metrics such as precision, recall, and F1
+score might be more informative in the case of imbalanced classes.
 
 ```{r}
+#| label: tbl-model-metrics
+#| tbl-cap: Confusion matrix and Accuracy measurements for model
 #| eval: true
@@
 #| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-# 
-# pp_check(first_model) +
-#   theme_classic() +
-#   theme(legend.position = "bottom")
-# 
-# posterior_vs_prior(first_model) +
-#   theme_minimal() +
-#   scale_color_brewer(palette = "Set1") +
-#   theme(legend.position = "bottom") +
-#   coord_flip()
+#| tbl-pos: 'H'
+#| fig-align: center
+
+# Print confusion matrix
+print(confusion_matrix)
+
+# Calculate accuracy
+accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
+
+# Print formatted accuracy with two decimal places
+cat("\nModel Accuracy: ", sprintf("%.2f", accuracy), "\n", sep = "")
 ```
 
-## Diagnostics
+## Model diagnostics
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+**Fit Diagnostics**:
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+The mean posterior predictive distribution (mean_PPD) shown in @tbl-ppd is the mean of the posterior distribution of the predicted probabilities for Pakistan winning a match.
+The value of 0.4 indicates that, on average, the model predicts Pakistan has a 40% chance of winning a match based on the predictors.
+
+**MCMC Diagnostics**:
+
+@tbl-mcmc shows diagnostic measurements of the sampling process to assess
+whether the model has converged and whether the samples are reliable.
+
+-   `mcse (Monte Carlo Standard Error)`: The MCSE indicates the standard
+    error of the Monte Carlo estimates for each parameter. Smaller
+    values are better, as they show that the estimates are stable and
+    not too variable. Most of the parameters here have small MCSE values
+    (close to 0), indicating stable estimates.
+
+-   `Rhat`: Rhat is a measure of convergence. An Rhat value of 1.0
+    indicates perfect convergence, meaning the chains for the MCMC
+    sampling are mixing well and that the estimates for each parameter
+    are reliable. All Rhat values here are exactly 1.0, which suggests
+    that the sampling chains converged properly.
+
+-   `n_eff (Effective Sample Size)`: This is a measure of how many
+    independent samples are available for each parameter. Larger values
+    indicate better sampling. Here, most parameters have high effective
+    sample sizes (ranging from \~1600 to \~3000), suggesting that the
+    model has gathered a sufficiently large and diverse set of samples
+    for each parameter, ensuring robust inference.
 
 ```{r}
+#| label: tbl-ppd
+#| tbl-cap: Model posterior predictive distribution
+#| eval: true
 #| echo: false
-#| eval: true
 #| message: false
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
 
-# plot(first_model, "trace")
-# 
-# plot(first_model, "rhat")
+model_summary <- summary(model)
+model_summary <- as.data.frame(model_summary)
+
+# Extract relevant fit diagnostics
+fit_diagnostics <- model_summary[c("mean_PPD"),c(1, 3:6)]
+fit_diagnostics$variable <- "mean_PPD"
+fit_diagnostics <- fit_diagnostics[, c("variable", "mean", "sd", "10%", "50%", "90%")]
+
+# fit_diagnostics |>
+#   tt()
+#   style_tt(j = 2:5, align = "r") |>   # Align columns 2 to 5 to the right
+#   format_tt(digits = 1, num_fmt = "decimal")  # Format numbers with 1 decimal
+
+# Use kableExtra for table styling
+fit_diagnostics_rounded <- round(fit_diagnostics[, 2:6], 2)
+
+fit_diagnostics_rounded |>
+  kable(format = "latex")
+
+```
+
+```{r}
+#| label: tbl-mcmc
+#| tbl-cap: Model MCMC statisics
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+
+mcmc_row_names <- c("(Intercept)", "year", "team2Bangladesh", "team2England", "team2India", "team2Ireland", "team2New Zealand", "team2South Africa", "team2Sri Lanka", "team2West Indies", "team2Zimbabwe", "toss_win")
+
+model_summary2 <- summary(model)
+model_summary2 <- as.data.frame(model_summary2)
+fit_diagnostics2 <- model_summary2[mcmc_row_names,c(2, 7:8)]
+fit_diagnostics2$variable <- mcmc_row_names
+fit_diagnostics2 <- fit_diagnostics2[, c("variable", "mcse", "Rhat", "n_eff")]
+
+# fit_diagnostics2 |>
+#   tt() |>
+#   style_tt(j = 2:4, align = "r") |>
+#   format_tt(digits = 1, num_fmt = "decimal")
+
+# Use kableExtra for table styling
+fit_diagnostics_rounded2 <- round(fit_diagnostics2[, 2:4], 2)
+
+fit_diagnostics_rounded2 |>
+  kable(format = "latex")
 ```

### 2024-12-04 scripts/05-model_data.R

@@ -20,21 +20,10 @@ library(tidyr)
 #### Read data ####
 player_data <- read_parquet("data/02-analysis_data/cleaned_player_data.parquet")
 match_data <- read_parquet("data/02-analysis_data/cleaned_match_data.parquet")
-
-# Ensure Pakistan is always team1
-for (i in 1:nrow(match_data)) {
-  if (match_data$team1[i] == "Pakistan" || match_data$team2[i] == "Pakistan") {
-    if(match_data$team1[i] != "Pakistan"){
-      temp <- match_data$team1[i]
-      match_data$team1[i] <- "Pakistan"
-      match_data$team2[i] <- temp
-    }
-  }
-}
+pakistan_match_data <- read_parquet("data/02-analysis_data/cleaned_pakistan_match_data.parquet")
 
 # Filter match data for Pakistan
-pakistan_matches <- match_data %>%
-  filter(team1 == "Pakistan") %>% 
+pakistan_matches <- pakistan_match_data %>%
   mutate(pakistan_win = case_when(
     winner == "Pakistan" ~ 1,
     TRUE ~ 0))%>%
@@ -42,8 +31,6 @@ pakistan_matches <- match_data %>%
     toss_winner == "Pakistan" ~ 1,
     TRUE ~ 0))%>%
   mutate(year = lubridate::year(date))
-
-
 
 # Summarize match-level performance
 pakistan_match_summary <- pakistan_matches %>%
@@ -100,6 +87,7 @@ confusion_matrix_first <- table(
   Actual = match_test_data$pakistan_win,
   Predicted = match_test_data$predicted_class_first
 )
+
 confusion_matrix_second <- table(
   Actual = year_test_data$wins,
   Predicted = year_test_data$predicted_class_second
@@ -112,57 +100,9 @@ accuracy_second <- sum(diag(confusion_matrix_second)) / sum(confusion_matrix_sec
 print(paste("Accuracy:", round(accuracy_first, 2)))
 print(paste("Accuracy:", round(accuracy_second, 2)))
 
-# Model for predicting players career? Not sure if needed
-pakistan_players <- player_data %>%
-  filter(country == "Pakistan")
-
-# Summarize player performance
-pakistan_player_summary <- pakistan_players %>%
-  group_by(player, years_played = end-start) %>%
-  summarize(
-    total_runs = sum(runs, na.rm = TRUE),
-    total_wickets = sum(wickets, na.rm = TRUE),
-    matches = n()
-  )
-player_yearly_summary <- player_data %>%
[…]

### 2024-12-04 other/model_card/datasheet_template.qmd
Final submission - added discussion, model card, and appendix

---
title: "Model Card for Predicting Pakistan Cricket Match Outcomes"
author: 
  - Muhammad Abdullah Motasim
thanks: "Model can me found at model/win_chance.rds and scripts/05-model_data.R"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: pdf
number-sections: true
---

# Model Card for Predicting Pakistan Cricket Match Outcomes

## Model Details
- **Model Name**: Pakistan Cricket Match Outcome Predictor  
- **Algorithm**: Logistic Regression using Bayesian Generalized Linear Model (via `stan_glm`)  
- **Purpose**: To predict the likelihood of Pakistan winning a cricket match based on historical match data.  
- **Version**: 1.0  
- **Developed By**: Muhammad Abdullah Motasim  
- **Frameworks Used**: R (`rstanarm`, `dplyr`)

## Intended Use
- **Primary Application**:  
  - Predict match outcomes for Pakistan cricket matches.  
  - Analyze factors influencing match results, such as opponent team, toss outcomes, and time trends.  
- **Intended Users**: Cricket analysts, statisticians, fans, and sports strategists.  
- **Use Cases**:  
  - Enhancing cricket strategies by identifying key predictors of success.  
  - Historical performance analysis of Pakistan’s cricket team.

## Model Metrics
- **Training Data Accuracy**: Not explicitly calculated; the model was fit on 80% of the data split.  
- **Testing Data Accuracy**: 82% (rounded from 0.82).  
- **Confusion Matrix**:  
  - **True Positives**: Correctly predicted Pakistan wins.  
  - **True Negatives**: Correctly predicted losses.  
  - **Errors**: Misclassifications in either class.  
- **Model Coefficients**:  
  - **Year**: `0.03 ± 0.04` (indicates a minor increasing trend in win probability over time).  
  - **Opponent Teams**:  
    - Strongest predictors:  
      - **Bangladesh** (`3.58 ± 1.35`) indicates a higher probability of wins.  
      - **West Indies** (`2.20 ± 0.86`) also shows a strong win probability.  
      - **Zimbabwe** (`2.47 ± 1.51`).  
    - Moderate predictors: **England**, **Sri Lanka**.  
    - Weak/uncertain predictors: **India**, **South Africa**, **New Zealand**.  
  - **Toss Win**: `0.59 ± 0.43` (winning the toss slightly increases the probability of winning the match).

## Training Data
- **Source**:  
  - Player and match data files: `cleaned_player_data.parquet`, `cleaned_match_data.parquet`, and `cleaned_pakistan_match_data.parquet`.  
- **Structure**: Includes match-level variables such as winner, toss winner, opponent team, and match year.  
- **Size**: Split into training (80%) and testing (20%) datasets.  
- **Preprocessing**:  
  - Filtering for Pakistan matches.  
  - Encoding binary variables (e.g., match win/loss, toss win/loss).

## Ethical Considerations
- **Bias**: The model's coefficients reveal some bias towards opponent teams (e.g., higher probabilities against weaker teams like Bangladesh and Zimbabwe). This reflects historical trends but might perpetuate biases in cricket analysis.  
- **Data Limitations**:  
  - The dataset does not include contextual factors like weather, player fitness, or ground conditions, which could influence outcomes.  
  - Matches against less frequent opponents may lead to inflated coefficients (e.g., Ireland).  
- **Fairness**: The model does not account for changes in cricket rules or playing conditions over time.  
- **Transparency**: Coefficients and assumptions are clearly documented to ensure interpretability.

## Caveats and Recommendations
- **Caveats**:  
  - The model assumes linear relationships between predictors and outcomes, which may oversimplify complex dynamics in cricket matches.  
  - The data includes only matches where Pakistan participated, limiting generalizability to other teams.  
  - The "toss win" variable shows moderate predictive power, but its causal relationship with match outcomes is not guaranteed.  
- **Recommendations**:  
  - Include additional contextual features (e.g., match location, player statistics) for better predictions.  
  - Validate the model on more recent datasets to ensure reliability over time.  
  - Be cautious when using the model for critical decision-making, as it simplifies cricket's multifaceted nature.

## Acknowledgments
This model was inspired by Pakistan’s rich cricketing history and aims to provide insights into factors affecting their performance. The Bayesian approach ensures probabilistic interpretability and robustness in predictions.

## Contact Information
- **Developer**: Muhammad Abdullah Motasim  
- **Email**: abdullah.motasim@mail.utoronto.ca 
- **GitHub Repository**: https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data/tree/main

### 2024-12-04 paper/paper.qmd
Final submission - added discussion, model card, and appendix

@@
 ---
-title: "Analyzing the Pakistan Cricket Team's Performance (2002-2024)"
-subtitle: "My subtitle if needed"
+title: "Analyzing Factors Affecting Pakistan Cricket Team's Performance (1952-2024)"
+subtitle: "Opponent and Coin Tossplay Play Crucial Role in Deciding Outcome"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper analyzes Pakistan's cricket performance from 2002 to 2024, using a linear regression model to examine factors influencing match outcomes.
+We find that opponent strength and toss outcomes are significant predictors of Pakistan's success, highlighting both internal team dynamics and external variables.
+These findings provide crucial insights for coaches, analysts, and administrators, offering strategies to improve performance under varying conditions.
+Ultimately, our analysis contributes to a deeper understanding of the factors that drive Pakistan’s cricket success, with implications for teams worldwide."
 format: pdf
@@
 
-player_data <- read_parquet("data/02-analysis_data/cleaned_player_data.parquet")
+player_data <- read_parquet(here("data/02-analysis_data/cleaned_player_data.parquet"))
 
@@
 
-Pakistan’s cricket team holds a storied legacy, marked by remarkable
-highs and challenging lows. From their 1992 World Cup triumph to
-fluctuations in form over the decades, analyzing trends in performance
-offers insights into their journey and areas for strategic improvement.
-
-Despite an extensive focus on player statistics and match highlights,
-there remains a lack of systematic statistical analysis to evaluate
-long-term performance trends. This paper fills this gap by analyzing
-data from 2002 to 2024 to understand what factors contribute to
-Pakistan’s success in international cricket.
+Pakistan’s cricket team holds a storied legacy, marked by remarkable highs and challenging lows.
+From their 1992 World Cup triumph to fluctuations in form over the decades, analyzing trends in performance offers insights into their journey and areas for strategic improvement.
+Despite the wealth of player statistics and match highlights available, there has been a lack of systematic statistical analysis that evaluates long-term performance trends.
+This paper fills that gap by analyzing data from 2002 to 2024 to uncover the key factors that contribute to Pakistan’s success in international cricket.
 
 We employ a linear regression model to analyze match outcomes,
-considering variables like year, opponent, and toss outcomes. The focus
-is on identifying trends and predicting the probability of victory under
-varying conditions. Our findings not only contribute to the literature
-on sports analytics but also provide actionable insights for coaches,
-analysts, and cricket administrators not only in Pakistan, but around
-the world.
+considering variables like year, opponent, and toss outcomes.
+Our findings reveal several important insights.
+First, the strength of the opposing team emerged as a significant factor influencing Pakistan’s chances of winning.
+We also found that toss outcomes had a noteworthy impact on match results.
+This paper highlights the importance of both internal factors (such as team performance) and external factors (such as opponent strength) in shaping match outcomes.
+These insights provide strategic guidance for coaches, analysts, and cricket administrators, not only in Pakistan but also around the world, on how to prepare and perform under varying conditions.
 
@@
 ggplot(data, aes(x = pakistan_win)) +
-  geom_bar(fill = "gold") +
+  geom_bar(fill = "darkblue") +
   labs(x = "Outcome", y = "Count") +
@@
 #| fig-subcap: ["Sri Lanka","New Zealand", "Australia", "England", "South Africa", "West Indies", "Zimbabwe", "Bangladesh", "Ireland", "India"]
-#| layout-ncol: 2
-#| layout-nrow: 5
+#| layout-ncol: 3
+#| layout-nrow: 4
 #| echo: false
@@
 
-- `Sri Lanka` (Figure 3a): Pakistan's matches against Sri Lanka are fairly balanced, with a slightly higher frequency of losses compared to wins.
+- `Sri Lanka`: Pakistan's matches against Sri Lanka are fairly balanced, with a slightly higher frequency of losses compared to wins.
 Draws are the least frequent outcome, suggesting a competitive dynamic between the two teams.
 
-- `New Zealand` (Figure 3b): Matches against New Zealand reveal a similar trend, with losses being more frequent than wins.
+- `New Zealand`: Matches against New Zealand reveal a similar trend, with losses being more frequent than wins.
 The proportion of draws remains relatively low, indicating fewer instances of matches being evenly contested.
 
-- `England` (Figure 3c): Pakistan has a notably higher number of losses against England, with draws constituting the smallest category.
+- `England`: Pakistan has a notably higher number of losses against England, with draws constituting the smallest category.
 This indicates that England has historically been a challenging opponent for Pakistan.
 
-- `Australia` (Figure 3d): A significant disparity is observed in matches against Australia, where losses outnumber wins by a considerable margin.
+- `Australia`: A significant disparity is observed in matches against Australia, where losses outnumber wins by a considerable margin.
 This trend underscores the dominance of Australia in these encounters.
 
-- `South Africa` (Figure 3e): Matches against South Africa also show a higher proportion of losses compared to wins.
+- `South Africa`: Matches against South Africa also show a higher proportion of losses compared to wins.
 However, the gap between these two outcomes is less pronounced compared to matches against Australia.
 
-- `West Indies` (Figure 3f): Pakistan exhibits a stronger performance against West Indies, with wins slightly outnumbering losses.
+- `West Indies`: Pakistan exhibits a stronger performance against West Indies, with wins slightly outnumbering losses.
 Draws remain the least frequent outcome, emphasizing decisive results in these encounters.
 
-- `Zimbabwe` (Figure 3g): Against Zimbabwe, Pakistan demonstrates a clear dominance, with wins significantly outnumbering losses.
+- `Zimbabwe`: Against Zimbabwe, Pakistan demonstrates a clear dominance, with wins significantly outnumbering losses.
 Draws are nearly negligible, highlighting Pakistan's consistent success in these matches.
 
-- `Bangladesh` (Figure 3h): Similar to Zimbabwe, Pakistan has a markedly higher number of wins against Bangladesh, with losses and draws being infrequent.
+- `Bangladesh`: Similar to Zimbabwe, Pakistan has a markedly higher number of wins against Bangladesh, with losses and draws being infrequent.
 This dominance underscores Pakistan's strong track record against this opponent.
@@
 #| label: fig-pk-wins
-#| fig-cap: "Number of Pakistan's Wins Over the Years"
+#| fig-cap: "Number of Pakistan's wins over the years"
 #| echo: false
@@
   theme_minimal()
-
 ```
@@
 
+```{r}
+#| label: fig-career-end-runs-individual
+#| fig-cap: "Career batting runs for each player by year of retirement"
+#| echo: false
+#| eval: true
+#| warning: false
+
+# Filter out players whose batting innings are greater than their bowling innings
+filtered_data <- player_data %>%
+  filter(bat_innings <= bowl_innings)
+
+# Add a new column to calculate career duration (end_year - start_year)
+filtered_data <- filtered_data %>%
+  mutate(career_duration = end - start)
+
+# Aggregate data to get total runs for each year when players ended their career
+runs_by_year <- filtered_data %>%
+  group_by(end) %>%
+  summarise(total_runs = sum(bat_runs))
+
+# Plot individual data points of each player's career runs:
+ggplot(filtered_data, aes(x = end, y = bat_runs)) +
+  geom_point(color = "darkblue", size = 2) +
+  labs(
+    # title = "Career Batting Runs for Each Player by Year Career Ended",
+    x = "Year Career Ended",
+    y = "Career Batting Runs"
+  ) +
+  theme_minimal()
+
+```
+
+```{r}
+#| label: fig-career-end-runs-all
+#| fig-cap: "Total career batting runs for all players retiring in each year"
+#| echo: false
+#| eval: true
+#| warning: false
+
+# Filter out players whose batting innings are greater than their bowling innings
+filtered_data <- player_data %>%
+  filter(bat_innings <= bowl_innings)
+
+# Add a new column to calculate career duration (end_year - start_year)
+filtered_data <- filtered_data %>%
+  mutate(career_duration = end - start)
+
+# Aggregate data to get total runs for each year when players ended their career
+runs_by_year <- filtered_data %>%
+  group_by(end) %>%
+  summarise(total_runs = sum(bat_runs))
+
+# Plot the aggregated data: Total career runs per year
+ggplot(runs_by_year, aes(x = end, y = total_runs)) +
+  geom_point(color = "darkblue", size = 3) + 
+  labs(
+    # title = "Total Career Batting Runs by Year Career Ended",
+    x = "Year Career Ended",
+    y = "Sum of Total Career Batting Runs"
+  ) +
+  theme_minimal()
+
+```
+
+@fig-career-end-runs-individual presents a scatter plot showing the years in which players retired and the number of runs they scored throughout their careers, with each dot representing an individual Pakistani player.
+The plot reveals that most players retired with fewer than 1,000 career runs, and there is generally no strong positive trend across the data.
+However, there is a notable exception around 2015, where we see a slight uptick in the number of runs.
+This increase is short-lived, as the trend declines again shortly after, until 2024, when a large number of players retired, many of whom had accumulated over 1,000 career runs.
+This scatter plot provides valuable insight into individual player performances, which can be used to explore broader trends in the data.
+
+@fig-career-end-runs-all shows a similar scatter plot, but in this case, each dot represents the total combined runs of all players retiring in a given year, with the y-axis now reflecting the sum of career runs for all retiring players in that year.
+Upon examining this graph, we observe some outliers, particularly in 2006 and 2024.
+Referring back to @fig-career-end-runs-individual, the 2006 outlier can be explained by a player retiring with around 6,800 runs.
+The 2024 outlier is due to the unusually high number of players retiring that year, including an exceptional player who ended their career with approximately 7,500 runs.
+Aside from these outliers, the overall trend remains relatively stagnant, with a slight increase again around 2015, though this quickly declines thereafter.
+
 # Discussion {#sec-discussion}
 
-## First discussion point {#sec-first-point}
+## Analysis Overview
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The
-discussion is a chance to show off what you know and what you learnt
-from all this.
+In this paper, I employed Bayesian Logistic Regression to estimate the probability of Pakistan winning cricket matches, considering key factors such as the year of the match, the opponent, and the toss outcome.
+The model was specifically designed to understand how these variables influence Pakistan's chances of winning, while accounting for uncertainty and variability in the data.
+To ensure the analysis focused on relevant information, the data was cleaned and processed to include only Pakistan’s matches, along with important player career statistics that could impact match outcomes.
+By incorporating these factors, the model provided a more comprehensive understanding of Pakistan’s performance dynamics.
 
-## Second discussion point
+The results of the analysis reveal valuable insights into Pakistan's performance trends over time.
+For example, the strength of the opposing team played a significant role, with match outcomes showing a clear correlation between Pakistan's success and the competitive level of the opposition.
+Additionally, the toss outcome also emerged as influential factors, providing a deeper understanding of how certain conditions affect Pakistan's likelihood of winning.
+These findings underscore the importance of both internal (team performance) and external (opponent strength, match context) factors in shaping match outcomes, offering strategic insights for future team preparation and performance analysis.
 
-Please don't use these as sub-heading labels - change them to be what
-your point actually is.
+## Effect of Opponent
 
-## Third discussion point
+One key finding from the analysis is the significant influence of the opposing team on Pakistan's match performance.
+As shown in @fig-model-results and @tbl-model-results, the calculated coefficient values for each opposing team align with the patterns observed in @fig-opponent-stats, where Pakistan has a higher success rate against teams with greater positive coefficient values, such as Bangladesh, and a lower success rate against teams with lower or negative coefficient values, such as South Africa.
+This correlation suggests that Pakistan's chances of winning are not only influenced by its own team performance but also by the strength of the opposition.
 
-## Weaknesses and next steps
+The model further emphasizes the variations in Pakistan's odds of winning depending on the opponent.
+For example, the coefficients for teams like Bangladesh, England, and Sri Lanka indicate more favorable matchups for Pakistan, suggesting that these teams have historically been more susceptible to Pakistan's strengths or less competitive overall.
+Conversely, the coefficients for teams like India and South Africa point to tougher challenges, with Pakistan's chances of winning being lower in these matchups.
+This pattern highlights the broader strategic landscape, where understanding the relative strength of the opponent becomes crucial for predicting outcomes.
+It underscores the need for teams to account not just for internal factors, such as player form or match conditions, but also for the competitive level of the opposition when preparing for matches.
+The results validate the intuition that the quality of the opposing team is a significant determinant of Pakistan’s performance, and that adjusting tactics accordingly can improve chances against stronger teams.
+Ultimately, the analysis reinforces the importance of a nuanced, strategic approach to match preparation, where the nature of the opposition is carefully considered alongside team strengths and other match variables.
 
-Weaknesses and next steps should also be included.
+## Matches Won Over the years
+
+When analyzing @fig-pk-wins in relation to @fig-career-end-runs-individual and @fig-career-end-runs-all, the trends observed align with what we would expect based on the historical performance of Pakistan’s cricket team.
+From around 2015, we see a noticeable increase in the number of games won by Pakistan.
+This can be explained by the presence of a number of veteran players who, with their vast experience, played a pivotal role in guiding the team to more victories.
+Around 2015, several seasoned players reached the peak of their careers or retired, bringing with them not only impressive career statistics but also crucial leadership and expertise on the field.
+The presence of these experienced players undoubtedly helped stabilize the team's performance and contributed to the higher win rates during this period.
+
+However, following 2015, there is a visible decline in Pakistan’s performance, with a fluctuating number of wins per year.
+This drop in consistent victories is likely due to the retirement of key veteran players who had been essential to Pakistan’s success.
+With their departure, a significant number of new, less experienced players entered the team.
+While these new players showed potential and skill, they lacked the depth of experience necessary to handle the pressures and complexities of international cricket.
+This transition phase, where fresh talent was integrated into the team, led to inconsistency in match outcomes, which we observe as fluctuations in the win data post-2015.
+
+Looking ahead to 2024, the retirement of another set of senior players further raises concerns about the future trajectory of Pakistan's performance.
+This continuous cycle of veteran players retiring and new players coming in without the same level of experience poses a challenge for the team.
+Although these new players bring in raw talent, the absence of experienced leaders and match-winners could contribute to a potential decline in overall performance.
+Given the historical patterns, we can reasonably anticipate a dip in the team's performance in the coming years unless the emerging players quickly adapt and gain the necessary experience to compete at the highest level.
+Therefore, while the retirement of key players is a natural part of the cycle, it highlights the ongoing challenge Pakistan faces in maintaining a competitive edge in international cricket.
+
+## Weaknesses and Limitations:
+
+Despite the valuable insights gained, there are several limitations in the data and methodology that must be acknowledged.
+One significant challenge was the limited number of matches involving Pakistan that could be used to train the model.
+The data set included only 820 match observations for Pakistan, which, while informative, is relatively small when considering the complexity of cricket as a sport with numerous variables influencing match outcomes.
+This limited sample size could affect the robustness of the model’s estimates, and in particular, the ability to generalize predictions across different conditions or time periods.
+
+Additionally, there was a lack of detailed match statistics in the raw data, which constrained the scope of the analysis.
+For instance, while data on match outcomes (win/loss), toss results, and opponents were available, detailed statistics like the number of runs scored, wickets taken, or batting and bowling performances were only provided for the winning team.
+This lack of comprehensive match data—especially for losing teams—meant that I could not include key performance indicators like runs or wickets in the model, as many of these entries were missing or incomplete.
+These missing values could have influenced the accuracy of the model, and it would be prudent to consider these missing data points in future iterations of the analysis, potentially employing imputation methods or seeking alternative data sources.
+
+Lastly, while the Bayesian Logistic Regression model is powerful, it assumes that the relationships between the predictors and the outcome are linear in nature.
+Although this is a reasonable assumption in many cases, cricket is a dynamic sport with complex interactions, and this model may not fully capture non-linear relationships or interactions between factors such as team composition, player injuries, or the impact of different match venues.
+
+## Next Steps
+
+Despite these limitations, there are several directions for future work that could improve the analysis and expand our understanding of Pakistan’s cricket performance.
+
+- Improving the Data Set: One of the key areas for future research would be to gather a more extensive data set, particularly focusing on adding detailed match statistics, including runs, wickets, and player performances for both teams.
+This would allow for a more comprehensive model, potentially including these factors as predictors in a multivariate analysis.
+Additionally, expanding the sample size to include more matches across a broader time period could help provide more robust and generalizable results.
+
+- Incorporating Player-Specific Metrics: Another avenue for improvement would be incorporating player-specific statistics into the model.
+By analyzing how the performance of individual players—batting and bowling averages, player form, and injury history—affects Pakistan’s match outcomes, the model could be enhanced to offer a more granular understanding of what drives success or failure in cricket matches.
+Including player career data alongside match data would offer a deeper, more detailed look into the team's performance and could be used to identify potential strengths and weaknesses.
+
+- Considering Interaction Effects: Future work could explore more advanced models that account for interactions between variables.
+For instance, the relationship between Pakistan’s performance and the opponent could vary depending on the specific conditions of a given year, such as the emergence of new players, changes in team strategy, or alterations in the competitive landscape of international cricket.
+Exploring such interaction effects could provide a more nuanced view of Pakistan's performance over time.
 
@@
 
-# Additional data details
+# Surveys, Sampling, and Observational Data
+
+In this paper, we focus on a dataset of cricket matches involving Pakistan and their performance outcomes.
+However, a significant challenge in our analysis was the limited and often incomplete data available for each match, which made it difficult to fully answer the question of what affects Pakistan’s overall match outcomes.
+Specifically, the dataset was missing detailed match statistics, such as the number of wickets per over, which player bowled in each over, individual player statistics, and detailed match events like runs and outs for both teams.
+This lack of granular data posed a substantial limitation to the analysis and highlighted the importance of accurate and complete data collection in sports analytics.
+
+## Data Collection
+In an ideal scenario, the data would have been far more comprehensive, particularly with respect to the granular details of each match.
+To overcome the limitations of the current dataset, it would be essential to gather detailed match data through a combination of observational data and video analysis.
+This would involve either live or recorded footage of each match, from which key statistics could be manually or automatically extracted.
+Such an approach would include the following steps:
+
+- `Detailed Over-by-Over Data`: Analyzing each over of the match would provide crucial insights into the match's dynamics.
+This would involve recording key statistics, such as the number of runs scored, the number of wickets taken, the bowler and batter in each over, and any significant events, like boundary hits or wickets.
+This granular data would help in identifying the ebb and flow of the match and how different events influence Pakistan’s chances of winning.
+
+- `Player-Specific Statistics`: In addition to match-level statistics, it is essential to record individual player data.
+For every player involved, detailed statistics such as runs scored, wickets taken, the number of innings played, and the number of outs would be recorded.
+This information would allow for a more precise understanding of how the performance of individual players influences the match outcome.
+For instance, if certain players consistently score high or take wickets in critical moments, this could provide valuable insight into the team’s overall performance strategy.
+
+- `Team-Specific Data`: Beyond individual player performance, it would be essential to capture team-level data for both the batting and bowling sides.
+This would include the number of runs scored by each team, the number of outs, and key tactical decisions like the batting order, field placements, and bowling changes.
+Understanding these aspects could allow for a better analysis of how team strategies affect the outcome of the match.
+
+## Annual Player Performance
+Another critical aspect that could improve the analysis is the collection of annual player performance data.
+While the current dataset allows for some insight into match outcomes, it lacks detailed player statistics across multiple years.
+To address this, it would be helpful to build a dataset that combines match-level data for each player over the course of several years.
+
+- `Annual Player Statistics`: For each player, aggregating data across multiple matches would provide a clearer picture of their performance over time.
+Key metrics such as the total number of runs scored, the number of wickets taken, the number of matches played, and the number of innings would be compiled.
+This would allow for a more nuanced analysis of individual player performance, considering factors like form and consistency.
+
+- `Linking Player Performance to Match Outcomes`: By examining the link between individual player statistics and match outcomes, it would be possible to identify trends that might not be evident at the match level.
+For example, players with higher annual runs might be correlated with higher chances of winning for the team, or certain bowlers who take more wickets in specific conditions could become key to success.
+This approach would allow for more precise modeling of match outcomes based on both individual and team performance.
+
+- `Player and Opponent Comparisons`: To better understand how player performance interacts with match outcomes, it would be useful to collect data on player performances not just for Pakistan, but also for opposing teams.
+By comparing the performances of Pakistan’s players with those of the opposing team’s players, we could gain insights into how the relative strength of players affects the result.
+For example, if Pakistan’s key players perform well but the opposing team’s star players outperform them, it could explain a loss despite good individual performances.
 
@@
 
-# fit_diagnostics |>
-#   tt()
+# table1 <- fit_diagnostics |>
+#   tt()|>
 #   style_tt(j = 2:5, align = "r") |>   # Align columns 2 to 5 to the right
-#   format_tt(digits = 1, num_fmt = "decimal")  # Format numbers with 1 decimal
+#   format_tt(digits = 1, num_fmt = "decimal")
 

### 2024-12-04 scripts/05-model_data.R

@@ -47,10 +47,6 @@ match_train_index <- createDataPartition(pakistan_matches$pakistan_win, p = 0.8,
 match_train_data <- pakistan_matches[match_train_index, ]
 match_test_data <- pakistan_matches[-match_train_index, ]
 
-year_train_index <- createDataPartition(pakistan_match_summary$wins, p = 0.8, list = FALSE)
-year_train_data <- pakistan_match_summary[year_train_index, ]
-year_test_data <- pakistan_match_summary[-year_train_index, ]
-
 first_model <-
   stan_glm(
     formula = pakistan_win ~  year + team2 + toss_win,
@@ -60,49 +56,25 @@ first_model <-
     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_aux = exponential(rate = 1, autoscale = TRUE),
   )
-
-second_model <- stan_glm(
-  formula = wins ~ year + total_matches + tosses_won,  # Predict the number of wins
-  data = year_train_data,  # Your training data
-  family = poisson(link = "log"),  # Use Poisson distribution for count data
-  prior = normal(0, 2.5),  # Prior for coefficients
-  prior_intercept = normal(0, 5),  # Prior for intercept
-  chains = 4,  # Number of chains for MCMC sampling
-  iter = 2000  # Number of iterations
-)
   
 summary(first_model)
-summary(second_model)
 
 #Predict probabilities for test data
 match_test_data$predicted_prob_first <- predict(first_model, newdata = match_test_data, type = "response")
-year_test_data$predicted_prob_second <- predict(second_model, newdata = year_test_data, type = "response")
 
 # Generate predicted classes based on a threshold (e.g., 0.5)
 match_test_data$predicted_class_first <- ifelse(match_test_data$predicted_prob_first > 0.5, 1, 0)
-year_test_data$predicted_class_second <- ifelse(year_test_data$predicted_prob_second > 0.5, 1, 0)
-
 # Evaluate model performance
 confusion_matrix_first <- table(
   Actual = match_test_data$pakistan_win,
   Predicted = match_test_data$predicted_class_first
 )
 
-confusion_matrix_second <- table(
-  Actual = year_test_data$wins,
-  Predicted = year_test_data$predicted_class_second
-)
 print(confusion_matrix_first)
-print(confusion_matrix_second)
 
[…]

### 2024-12-10 scripts/00-simulate_data.R
Updated simulating and testing simulation script to include match_data and test match_data

@@ -4,25 +4,39 @@
 # Date: 22 November 2024
 # Contact: abdullah.motasim@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: arrow, tibble, and readr libraries
+# Pre-requisites: dplyr arrow, and tibble libraries
 # Any other information needed? None
 
 
 #### Workspace setup ####
 library(tibble)
-library(readr)
+library(dplyr)
 library(arrow)
 set.seed(853)
 
 
 #### Simulate data ####
+players <- character(150)
 for (i in 1:150) {
   players[i] <- paste0("Player ", i)
 }
 
 
 # Countries
-countries <- c("India", "Australia", "England", "Pakistan", "South Africa", "Sri Lanka", "New Zealand", "West Indies", "Bangladesh", "Zimbabwe", "Afghanistan", "Ireland")
+countries <- c(
+  "India",
+  "Australia",
+  "England",
+  "Pakistan",
+  "South Africa",
+  "Sri Lanka",
+  "New Zealand",
+  "West Indies",
+  "Bangladesh",
+  "Zimbabwe",
+  "Afghanistan",
+  "Ireland"
+)
 
 # Simulated cricket data
 simulated_player_data <- tibble(
@@ -33,36 +47,110 @@ simulated_player_data <- tibble(
     replace = TRUE,
     prob = c(0.2, 0.15, 0.15, 0.1, 0.1, 0.1, 0.1, 0.05, 0.025, 0.025, 0.02, 0.02) # Rough country distribution
   ),
-  Start = sample(1970:2023, size = 150, replace = TRUE), # Random start years
-  End = sample(1985:2023, size = 150, replace = TRUE), # Random end years
-  Matches = sample(1:500, size = 150, replace = TRUE), # Number of matches played
-  Innings = sample(1:500, size = 150, replace = TRUE), # Number of innings
-  NotOuts = sample(0:50, size = 150, replace = TRUE), # Number of not-outs
-  BattingRunsScored = sample(0:20000, size = 150, replace = TRUE), # Total runs scored
-  HighScore = sample(50:400, size = 150, replace = TRUE), # Highest individual score
-  Average = round(runif(150, 20, 60), 2), # Batting average
-  BallsFaced = sample(1000:30000, size = 150, replace = TRUE), # Balls faced
-  StrikeRate = round(runif(150, 50, 150), 2), # Strike rate
-  Hundreds = sample(0:100, size = 150, replace = TRUE), # Number of centuries
[…]

### 2024-12-10 scripts/01-test_simulated_data.R

@@ -1,117 +1,135 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated data 
+# Purpose: Tests the structure and validity of the simulated data
 # Author: Muhammad Abdullah Motasim
 # Date: 22 November 2024
 # Contact: abdullah.motasim@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: arrow and tidyverse libraries
+# Pre-requisites: tidyverse, testthat, and lubridate libraries
 # Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
+library(testthat)
+library(lubridate)
 
-simulated_data <- read_parquet("data/00-simulated_data/simulated_player_data.parquet")
+# Load the simulated datasets
+match_data <- read_parquet("data/00-simulated_data/simulated_match_data.parquet")
+player_data <- read_parquet("data/00-simulated_data/simulated_player_data.parquet")
 
-# Test if the data was successfully loaded
-if (exists("simulated_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
+#### Tests for simulated_match_data ####
 
+# Test that there are 100 matches
+test_that("simulated_match_data contains 100 matches", {
+  expect_equal(nrow(match_data), 100)
+})
 
-#### Test data ####
+# Test that match_data has 10 columns
+test_that("simulated_match_data has 10 columns", {
+  expect_equal(ncol(match_data), 10)
+})
 
-# Check if the dataset has 150 rows
-if (nrow(simulated_data) == 150) {
-  message("Test Passed: The dataset has 150 rows.")
-} else {
-  stop("Test Failed: The dataset does not have 150 rows.")
-}
+# Test that match_data columns are of the correct type
+test_that("simulated_match_data column types are as expected", {
+  expect_type(match_data$team1, "character")
+  expect_type(match_data$team2, "character")
+  expect_true(lubridate::is.Date(match_data$date)) # Check if 'date' is a lubridate Date
+  expect_type(match_data$winner, "character")
+  expect_type(match_data$team1_runs, "integer") # Numeric columns
+  expect_type(match_data$team2_runs, "integer")
+  expect_type(match_data$team1_wickets, "integer")
+  expect_type(match_data$team2_wickets, "integer")
+  expect_type(match_data$toss_winner, "character")
+})
 
-# Check if the dataset has the expected number of columns
[…]

### 2024-12-10 scripts/02-download_data.R
Updated commenting and formatting for scripts 02-05

@@ -11,10 +11,12 @@
 #### Workspace setup ####
 library(cricketdata)
 library(tidyverse)
-library(dplyr) # For data wrangling
+library(dplyr)
 
 #### Download data ####
 
+# Fetch cricket player data from Cricinfo for various skill categories (batting, bowling, fielding)
+# for different formats (Test, ODI, T20). For each category, we fetch data for male players in career statistics.
 cricinfo_data1 <- fetch_cricinfo("test", "men", "batting", "career", country = NULL)
 cricinfo_data2 <- fetch_cricinfo("test", "men", "bowling", "career", country = NULL)
 cricinfo_data3 <- fetch_cricinfo("test", "men", "fielding", "career", country = NULL)
@@ -25,7 +27,7 @@ cricinfo_data7 <- fetch_cricinfo("t20", "men", "batting", "career", country = NU
 cricinfo_data8 <- fetch_cricinfo("t20", "men", "bowling", "career", country = NULL)
 cricinfo_data9 <- fetch_cricinfo("t20", "men", "fielding", "career", country = NULL)
 
-# Combine datasets by skill type (batting, bowling, fielding)
+# Data cleaning for the batting datasets
 cricinfo_data1 <- cricinfo_data1 %>%
   mutate(Player = str_trim(Player)) %>% # Remove leading/trailing spaces
   mutate(Player = str_to_title(Player)) # Ensure consistent capitalization
@@ -36,34 +38,36 @@ cricinfo_data7 <- cricinfo_data7 %>%
   mutate(Player = str_trim(Player)) %>%
   mutate(Player = str_to_title(Player))
 
+# Remove duplicates from the batting datasets
 cricinfo_data1 <- cricinfo_data1 %>% distinct()
 cricinfo_data4 <- cricinfo_data4 %>% distinct()
 cricinfo_data7 <- cricinfo_data7 %>% distinct()
 
-
+# Combine the datasets for batting data across all formats (Test, ODI, T20)
 combined_bat_data <- bind_rows(cricinfo_data1, cricinfo_data4, cricinfo_data7) %>%
   group_by(Player) %>%
   summarise(
     Country = first(Country), # Assuming 'Country' remains consistent for each player
-    Start = min(Start, na.rm = TRUE), # Take the earliest Start year
-    End = max(End, na.rm = TRUE), # Take the latest End year
[…]

### 2024-12-10 scripts/03-clean_data.R

@@ -14,6 +14,7 @@ library(dplyr)
 library(arrow)
 
 #### Clean data ####
+# Read raw cricket data files into R from CSV files located in the 01-raw_data folder.
 raw_player_data <- read_csv("data/01-raw_data/cricsheet_player_data.csv")
 raw_match_data <- read_csv("data/01-raw_data/cricsheet_match_data.csv")
 raw_meta_data <- read_csv("data/01-raw_data/player_meta_data.csv")
@@ -21,35 +22,43 @@ raw_bat_data <- read_csv("data/01-raw_data/combined_bat_data.csv")
 raw_bowl_data <- read_csv("data/01-raw_data/combined_bowl_data.csv")
 raw_field_data <- read_csv("data/01-raw_data/combined_field_data.csv")
 
+# Clean and reshape the batting data
 raw_bat_data <- raw_bat_data %>%
   mutate(bat_innings = Innings, bat_runs = Runs) %>%
   select(-Innings, -Runs)
 
+# Clean and reshape the bowling data
 raw_bowl_data <- raw_bowl_data %>%
   mutate(bowl_innings = Innings, bowl_runs = Runs) %>%
   select(-Innings, -Runs)
 
+# Clean and reshape the fielding data
 raw_field_data <- raw_field_data %>%
   mutate(field_innings = Innings) %>%
   select(-Innings)
 
+# Clean match data: Convert column names to lowercase, select relevant columns, and clean date format.
 cleaned_match_data <-
   raw_match_data |>
   janitor::clean_names() |>
-  select(team1,
-         team2,
-         date,
-         winner,
-         winner_wickets,
-         winner_runs,
-         toss_winner) |>
+  select(
+    team1,
+    team2,
+    date,
+    winner,
+    winner_wickets,
+    winner_runs,
+    toss_winner
+  ) |>
   mutate(date = as.character(date)) |>
   mutate(date = ymd(date)) |>
-  mutate(winner = ifelse(is.na(winner), "Draw", winner))  # Replace NA with "Draw" in winner column
+  mutate(winner = ifelse(is.na(winner), "Draw", winner)) # Replace NA with "Draw" in winner column
 
+# Correct the order of teams, ensuring Pakistan is always 'team1' if involved in the match.
 for (i in 1:nrow(cleaned_match_data)) {
   if (cleaned_match_data$team1[i] == "Pakistan" || cleaned_match_data$team2[i] == "Pakistan") {
-    if(cleaned_match_data$team1[i] != "Pakistan"){
[…]

### 2024-12-10 scripts/04-test_analysis_data.R

@@ -19,86 +19,95 @@ player_data <- read_parquet("data/02-analysis_data/cleaned_player_data.parquet")
 
 #### Test data ####
 
-# Tests for `cleaned_match_data
+# Tests for `cleaned_match_data`
 
+# Test to check if 'cleaned_match_data' has exactly 7 columns.
 test_that("cleaned_match_data 7 columns", {
   expect_equal(ncol(match_data), 7)
 })
 
+# Test to ensure that the column types in 'cleaned_match_data' are as expected.
 test_that("cleaned_match_data column types are as expected", {
-  expect_type(match_data$team1, "character")
-  expect_type(match_data$team2, "character")
-  expect_true(lubridate::is.Date(match_data$date))  # Check if 'date' is a lubridate Date
-  expect_type(match_data$winner, "character")
-  expect_type(match_data$winner_wickets, "double") # Numeric columns
-  expect_type(match_data$winner_runs, "double")
-  expect_type(match_data$toss_winner, "character")
+  expect_type(match_data$team1, "character") # 'team1' should be a character (text) column
+  expect_type(match_data$team2, "character") # 'team2' should be a character (text) column
+  expect_true(lubridate::is.Date(match_data$date)) # 'date' should be a Date type, validated using lubridate
+  expect_type(match_data$winner, "character") # 'winner' should be a character (text) column
+  expect_type(match_data$winner_wickets, "double") # 'winner_wickets' should be a numeric (double) column
+  expect_type(match_data$winner_runs, "double") # 'winner_runs' should be a numeric (double) column
+  expect_type(match_data$toss_winner, "character") # 'toss_winner' should be a character (text) column
 })
 
-# Test that `winner_wickets` and `winner_runs` columns in `cleaned_match_data` have at least one NA value (indicating mutually exclusive fields)
+# Test to ensure that either 'winner_wickets' or 'winner_runs' is NA for each row, indicating that only one can have a value (mutually exclusive).
 test_that("cleaned_match_data has mutually exclusive winner_wickets and winner_runs", {
-  expect_true(all(is.na(match_data$winner_wickets) | is.na(match_data$winner_runs)))
+  expect_true(all(is.na(match_data$winner_wickets) | is.na(match_data$winner_runs))) # Ensure only one column has a value at a time
 })
 
-# Test that `cleaned_match_data` has no duplicated rows
+# Test to check if there are any duplicate rows in 'cleaned_match_data'.
[…]

### 2024-12-10 scripts/05-model_data.R

@@ -5,7 +5,7 @@
 # Contact: abdullah.motasim@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: tidyverse, rstanarm, caret, dplyr, tidyr, and arrow libraries
-# Any other information needed?  The cleaned match and player data should be saved 
+# Any other information needed?  The cleaned match and player data should be saved
 # in 02-analysis_data folder
 
 
@@ -14,52 +14,52 @@ library(tidyverse)
 library(rstanarm)
 library(arrow)
 library(caret) # For data splitting and evaluation
-library(dplyr)
-library(tidyr)
 
 #### Read data ####
 player_data <- read_parquet("data/02-analysis_data/cleaned_player_data.parquet")
 match_data <- read_parquet("data/02-analysis_data/cleaned_match_data.parquet")
 pakistan_match_data <- read_parquet("data/02-analysis_data/cleaned_pakistan_match_data.parquet")
 
-# Filter match data for Pakistan
+# Add binary column for Pakistan won match to train model
 pakistan_matches <- pakistan_match_data %>%
   mutate(pakistan_win = case_when(
     winner == "Pakistan" ~ 1,
-    TRUE ~ 0))%>%
+    TRUE ~ 0
+  )) %>%
   mutate(toss_win = case_when(
     toss_winner == "Pakistan" ~ 1,
-    TRUE ~ 0))%>%
-  mutate(year = lubridate::year(date))
+    TRUE ~ 0
+  )) %>%
+  mutate(year = lubridate::year(date)) # Extract year from date column into seperate column
 
 # Summarize match-level performance
 pakistan_match_summary <- pakistan_matches %>%
-  group_by(year = lubridate::year(date)) %>%
+  group_by(year = lubridate::year(date)) %>% # Group data by the year of the match
   summarize(
-    total_matches = n(),
-    wins = sum(pakistan_win),
-    win_rate = mean(pakistan_win),
-    tosses_won = sum(toss_win)  # Sum the toss_win column to count tosses won by Pakistan
+    total_matches = n(), # Count the total number of matches in each year
+    wins = sum(pakistan_win), # Sum the 'pakistan_win' column to get total wins in each year
+    win_rate = mean(pakistan_win), # Calculate the win rate as the mean of 'pakistan_win'
+    tosses_won = sum(toss_win) # Sum the 'toss_win' column to count tosses won by Pakistan
   )
 
 set.seed(853)
[…]

### 2024-12-11 other/model_card/datasheet_template.qmd
Updated model card to include relevant references

@@
   - Muhammad Abdullah Motasim
-thanks: "Model can me found at model/win_chance.rds and scripts/05-model_data.R"
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
+thanks: "Model can me found at https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data/tree/main/models"
+date: today
+date-format: long
 format: pdf
 number-sections: true
+bibliography: references.bib
 ---
 
-# Model Card for Predicting Pakistan Cricket Match Outcomes
+# Model Details
+- **Developed By**: Muhammad Abdullah Motasim  
+- **Contact**: abdullah.motasim@mail.utoronto.ca
+- **License**: MIT
+- **Model Name**: Pakistan Cricket Match Outcome Predictor  
+- **Version**: 1.0  
+- **Algorithm**: Logistic regression using a Bayesian generalized linear model, trained on an 80-20 train-test split of data sourced from `cricketdata` [@cricketdata].
+- **Features**:  
+  - `Year` (numerical)  
+  - `Opponent Team` (categorical: `team2`)  
+  - `Toss Win` (binary: 1 for won, 0 for lost) 
+- **Purpose**: To predict the likelihood of Pakistan winning a cricket match based on historical match data.  
+- **Frameworks Used**: R [@citeR] with packages: `tidyverse` [@tidyverse], `rstanarm` [@rstanarm],  `arrow` [@arrow], `caret` [@caret])  
+- **Questions or Comments**: [abdullah.motasim@mail.utoronto.ca](mailto:abdullah.motasim@mail.utoronto.ca)
 
-## Model Details
-- **Model Name**: Pakistan Cricket Match Outcome Predictor  
-- **Algorithm**: Logistic Regression using Bayesian Generalized Linear Model (via `stan_glm`)  
-- **Purpose**: To predict the likelihood of Pakistan winning a cricket match based on historical match data.  
-- **Version**: 1.0  
-- **Developed By**: Muhammad Abdullah Motasim  
-- **Frameworks Used**: R (`rstanarm`, `dplyr`)
-
-## Intended Use
-- **Primary Application**:  
+# Intended Use
+- **Primary Intended Uses**:  
   - Predict match outcomes for Pakistan cricket matches.  
   - Analyze factors influencing match results, such as opponent team, toss outcomes, and time trends.  
-- **Intended Users**: Cricket analysts, statisticians, fans, and sports strategists.  
-- **Use Cases**:  
-  - Enhancing cricket strategies by identifying key predictors of success.  
-  - Historical performance analysis of Pakistan’s cricket team.
+- **Primary Intended Users**: Cricket analysts, statisticians, fans, and sports strategists.  
+- **Out-of-scope uses**:  
+  - Real-time decision-making or betting applications.
+  - Inferences about non-Pakistan matches or players.
 
-## Model Metrics
-- **Training Data Accuracy**: Not explicitly calculated; the model was fit on 80% of the data split.  
-- **Testing Data Accuracy**: 82% (rounded from 0.82).  
-- **Confusion Matrix**:  
-  - **True Positives**: Correctly predicted Pakistan wins.  
-  - **True Negatives**: Correctly predicted losses.  
-  - **Errors**: Misclassifications in either class.  
-- **Model Coefficients**:  
-  - **Year**: `0.03 ± 0.04` (indicates a minor increasing trend in win probability over time).  
-  - **Opponent Teams**:  
-    - Strongest predictors:  
-      - **Bangladesh** (`3.58 ± 1.35`) indicates a higher probability of wins.  
-      - **West Indies** (`2.20 ± 0.86`) also shows a strong win probability.  
-      - **Zimbabwe** (`2.47 ± 1.51`).  
-    - Moderate predictors: **England**, **Sri Lanka**.  
-    - Weak/uncertain predictors: **India**, **South Africa**, **New Zealand**.  
-  - **Toss Win**: `0.59 ± 0.43` (winning the toss slightly increases the probability of winning the match).
+# Factors
+## Groups
+- **Definition**: Groups refer to distinct categories represented in the evaluation data. For this model, the relevant groups are defined based on:  
+  - Opponent teams (`team2`).  
+  - Match year (`year`).  
+  - Toss win outcomes (`toss_win`).
 
-## Training Data
+- **Intersectional Analysis**: The model evaluates performance variations across combinations of year, opponent, and toss outcomes to assess fairness and representativeness.
+
+## Instrumentation
+- **Input Data Instruments**:  
+  - Match data sources include cricket archives and datasets standardized for analysis.  
+  - Input data preprocessing ensures consistent formats, removing noise in features.
+
+- **Potential Variability**: Accuracy might vary depending on the completeness and standardization of match records.
+
+## Environment
+- **Context of Deployment**: The model is intended for offline analysis and reporting.  
+- **Environmental Factors**: External influences (e.g., match conditions, venue) are not explicitly modeled but may implicitly affect historical data trends.
+
+## Relevant Factors  
+- **Definition**: Foreseeable salient factors for which model performance may vary include:  
+  - Opponent teams: Performance trends differ significantly based on the opposing team's skill level and strategies.  
+  - Match years: Historical shifts in team dynamics and cricket rules may influence predictions.  
+  - Toss outcomes: Toss-winning trends are crucial in determining match results.
+
+- **Determination**: These factors were identified based on cricket domain knowledge and statistical analysis of historical data trends.
+
+## Evaluation Factors  
+- **Definition**:  
+  - The evaluation reports include model performance across:  
+    - Opponent teams.  
+    - Toss win outcomes.  
+    - Yearly trends.
+
+- **Reasoning for Selection**:  
+  - These factors align with the model's intended use case: analyzing historical patterns and trends in cricket.  
+  - Relevant factors like player-specific attributes were excluded due to limited data availability.
+
+- **Relevance vs.
+Evaluation Factors**: While player-specific attributes or match venue conditions might be relevant for understanding outcomes, they were not included due to the lack of annotated datasets.
+Incorporating these factors would improve granularity but requires further data collection and standardization.
+
+# Model Metrics
+- **Model Performance Measures**:  
+  - Accuracy: 0.70.  
+  - Confusion matrix: Stored in `models/confusion_matrix_win_chance.rds`.  
+- **Decision Thresholds**:  
+  - Binary classification threshold of 0.5 for predicting wins.  
+- **Evaluation Variations**:  
+  - Cross-validation to analyze robustness.  
+  - Subset evaluation by year and opponent.
+
+# Evaluation Data
+- **Datasets**:  
+  - Cricket match data sourced from cricketdata R package, cleaned and aggregated into:  
+    - Player data (`data/02-analysis_data/cleaned_player_data.parquet`)  
+    - Match data (`data/02-analysis_data/cleaned_match_data.parquet`)  
+    - Pakistan-specific matches (`data/02-analysis_data/cleaned_pakistan_match_data.parquet`)  
+- **Motivation**: To understand performance patterns of Pakistan's cricket team over time and against different opponents.  
+- **Preprocessing**:  
+  - `Data Reshaping`: Renamed and reorganized columns in the batting, bowling, and fielding datasets to standardize the data (e.g., bat_innings, bat_runs) and removed unnecessary columns.
+
+  - `Match Data Cleaning`: Converted column names to lowercase, selected relevant columns, and cleaned date formats; replaced missing "winner" values with "Draw".
+
+  - `Team Reordering`: Ensured Pakistan was always listed as team1 in matches involving Pakistan, swapping team order when necessary.
+
+  - `Filtering`: Subsetted match data to include only matches where Pakistan was team1 and filtered player data to include only Pakistani players.
+
+  - `Data Aggregation`: Combined batting, bowling, and fielding data, then summarized player statistics (e.g., total innings, runs, wickets) while calculating batting averages, handling edge cases.
+
+# Training Data
 - **Source**:  
-  - Player and match data files: `cleaned_player_data.parquet`, `cleaned_match_data.parquet`, and `cleaned_pakistan_match_data.parquet`.  
+  - Player and match data files: 
+    - `cleaned_player_data.parquet`
+    - `cleaned_match_data.parquet`
+    - `cleaned_pakistan_match_data.parquet`.  
 - **Structure**: Includes match-level variables such as winner, toss winner, opponent team, and match year.  
@@
 - **Preprocessing**: 
-  - Filtering for Pakistan matches.  
-  - Encoding binary variables (e.g., match win/loss, toss win/loss).
+  - `Added Binary Columns`: Created two new binary columns (pakistan_win and toss_win), encoding whether Pakistan won the match and toss, respectively (1 for win, 0 for loss).
 
-## Ethical Considerations
+  - `Calculated Aggregates`: Summarized match-level performance by calculating total matches, total wins, win rate, and tosses won by Pakistan for each year.
+
+# Quantitative Analyses
+- **Unitary Results**:  
+  - Overall accuracy and metrics calculated for each year and opponent group.  
+- **Intersectional Results**:  
+  - Performance evaluation across intersections of toss win outcomes and opponents.
+
+# Ethical Considerations
 - **Bias**: The model's coefficients reveal some bias towards opponent teams (e.g., higher probabilities against weaker teams like Bangladesh and Zimbabwe).
@@
 - **Fairness**: The model does not account for changes in cricket rules or playing conditions over time.
-- **Transparency**: Coefficients and assumptions are clearly documented to ensure interpretability.
 
-## Caveats and Recommendations
+- **Use Case and Impact**: While the model is primarily designed for analyzing Pakistan’s historical match data, it could also be used to predict outcomes in future matches, influencing decisions in strategic planning, media analysis, or fan engagement.
+However, it should not be used for critical decision-making, such as betting, as its predictions are based on historical patterns and do not account for all real-time variables.
+
+- **Mitigation**: To mitigate the potential risks of bias and inaccuracies, it is recommended to:
+
+  - Use the model as a complementary tool rather than the sole decision-making basis.
+
+  - Incorporate additional contextual features (e.g., player statistics, match conditions) in future model iterations to improve accuracy and fairness.
+
+- **Potential Risks**:
+
+  - Misuse: There is a risk that the model could be used in ways not intended, such as in gambling or betting, where the consequences of incorrect predictions could have financial or reputational implications.
+
+  - Over-reliance: Relying too heavily on the model's predictions without considering external factors (e.g., player injuries or weather) could lead to misinformed decisions.
+
+# Caveats and Recommendations
 - **Caveats**:  
@@
   - The "toss win" variable shows moderate predictive power, but its causal relationship with match outcomes is not guaranteed.  
+  - Excludes critical factors like player performance and weather conditions.  
 - **Recommendations**:  
   - Include additional contextual features (e.g., match location, player statistics) for better predictions.  
-  - Validate the model on more recent datasets to ensure reliability over time.  
-  - Be cautious when using the model for critical decision-making, as it simplifies cricket's multifaceted nature.
+  - Use the model as a supplementary tool for research and analysis.  
+  - Regularly update the model with new data for better performance and relevance.
 
-## Acknowledgments
-This model was inspired by Pakistan’s rich cricketing history and aims to provide insights into factors affecting their performance.
-The Bayesian approach ensures probabilistic interpretability and robustness in predictions.
+\newpage
 
-## Contact Information
-- **Developer**: Muhammad Abdullah Motasim  
-- **Email**: abdullah.motasim@mail.utoronto.ca 
-- **GitHub Repository**: https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data/tree/main
+# References

### 2024-12-11 paper/paper.qmd
Updated paper to remove unneccesary words

@@
 title: "Analyzing Factors Affecting Pakistan Cricket Team's Performance (1952-2024)"
-subtitle: "Opponent and Coin Tossplay Play Crucial Role in Deciding Outcome"
+subtitle: "Opponent and Coin Tossplay Play Key Role in Deciding Outcome"
 author: 
@@
 We find that opponent strength and toss outcomes are significant predictors of Pakistan's success, highlighting both internal team dynamics and external variables.
-These findings provide crucial insights for coaches, analysts, and administrators, offering strategies to improve performance under varying conditions.
+These findings provide information for coaches, analysts, and administrators, offering strategies to improve performance under varying conditions.
 Ultimately, our analysis contributes to a deeper understanding of the factors that drive Pakistan’s cricket success, with implications for teams worldwide."
@@
 confusion_matrix <- readRDS(here("models/confusion_matrix_win_chance.rds"))
-
 ```
@@
 Pakistan’s cricket team holds a storied legacy, marked by remarkable highs and challenging lows.
-From their 1992 World Cup triumph to fluctuations in form over the decades, analyzing trends in performance offers insights into their journey and areas for strategic improvement.
+From their 1992 World Cup triumph to fluctuations in form over the decades, analyzing trends in performance offers information into their journey and areas for strategic improvement.
 Despite the wealth of player statistics and match highlights available, there has been a lack of systematic statistical analysis that evaluates long-term performance trends.
@@
 considering variables like year, opponent, and toss outcomes.
-Our findings reveal several important insights.
+Our findings reveal several important observations.
 First, the strength of the opposing team emerged as a significant factor influencing Pakistan’s chances of winning.
@@
 This paper highlights the importance of both internal factors (such as team performance) and external factors (such as opponent strength) in shaping match outcomes.
-These insights provide strategic guidance for coaches, analysts, and cricket administrators, not only in Pakistan but also around the world, on how to prepare and perform under varying conditions.
+These observations provide strategic guidance for coaches, analysts, and cricket administrators, not only in Pakistan but also around the world, on how to prepare and perform under varying conditions.
 
@@
 [@cricsheet]. It contains details on the individual match year, teams,
-result, and toss winner. Additionally, the data set features
-comprehensive career statistics for individual players, including total
+result, and toss winner. Additionally, the data set features career statistics for individual players, including total
 runs, total balls faced, total wickets, and more. This data set was
@@
 
-Figures @fig-match-stats displays the count of several important match
-statistics. From these, we can observe that Pakistan faces a variety of
-opponents, with the most frequent ones being England, Sri Lanka, and
-Australia. Notably, Pakistan has had a challenging record against
-Australia, losing a significant number of matches, while many of their
-matches ended in draws.
+Figures @fig-match-stats show key match statistics for Pakistan.
+The data shows that Pakistan frequently faces England, Sri Lanka, and Australia, with a particularly poor record against Australia, including several draws.
 
@@
 compared to other teams, which accounts for the higher statistics shown.
-These figures represent Pakistan's performance specifically, not that of
-other teams.
-
-The following two graphs show the number of matches and tosses that
-Pakistan won or lost. The win-loss ratio is approximately 0.5,
-indicating that Pakistan has lost about twice as many matches as it has
-won. Additionally, the coin toss win ratio is close to even, reflecting
+
+The two graphs showing the number of matches and tosses that
+Pakistan won or lost inform us that Pakistan's win-loss ratio is approximately 0.5.
+This indicates that Pakistan has lost about twice as many matches as it has won.
+Additionally, the coin toss win ratio is close to even, reflecting
 a near 50/50 chance of Pakistan winning the toss.
 
-Finally, the last graph illustrates the number of matches Pakistan
-played each year. We can see a peak in 2016, with a gradual decline in
-matches played over the subsequent years, culminating in a sharp drop in
-the most recent year—comparable to the period during the COVID-19
-pandemic.
+Finally, the graph illustrates the number of matches Pakistan
+played each year has a peak in 2016. It then gradually declines over the subsequent years, culminating in a sharp drop in the most recent year—comparable to the period during the COVID-19 pandemic.
 
@@
     winner == "Pakistan" ~ 1,
-    TRUE ~ 0))%>%
+    TRUE ~ 0
+  )) %>%
   mutate(toss_win = case_when(
     toss_winner == "Pakistan" ~ 1,
-    TRUE ~ 0))%>%
+    TRUE ~ 0
+  )) %>%
   mutate(year = lubridate::year(date))
@@
 
-The model incorporates three independent variables that may influence
-match outcomes: year, opponent, and toss winner. The dependent variable
-is match_outcome, where 1 indicates a win for Pakistan and 0 indicates a
-loss. This setup allows us to examine the influence of key variables on
-Pakistan's performance.
+The model was trained in R utilizing the rstanarm package and it incorporates three independent variables that may influence match outcomes: year, opponent, and toss winner.
+The dependent variable is match_outcome, where 1 indicates a win for Pakistan and 0 indicates a loss.
+This setup allows us to examine the influence of key variables on Pakistan's performance.
 
@@
 likelihood of winning. This approach helps uncover trends in Pakistan's
-performance over time and under varying conditions, providing insights
+performance over time and under varying conditions, providing intuition
 into factors that may inform future match strategies.
@@
 
--   `Prior for the Coefficients (year, team2, toss_win)`: Each of the
-    model coefficients is assigned a Normal prior with a mean of 0 and a
-    standard deviation of 2.5. This reflects the assumption that we
-    expect the effects of the predictors to be centered around zero,
-    with a reasonable amount of uncertainty (a wide spread due to the
-    scale of 2.5) about their true values.
-
--   `Prior for the Intercept`: A normal prior with the same parameters
-    as the coefficients (mean = 0, scale = 2.5) is used for the
-    intercept. The intercept represents the baseline log-odds of
-    Pakistan winning a match when all other predictors are zero.
-
--   `Prior for the Auxiliary Parameter (prior_aux)`: This parameter
-    represents the dispersion of the logistic regression model. An
-    Exponential prior with a rate of 1 is used, which is typical in
-    Bayesian models for logistic regression. The Exponential prior
-    allows for a distribution with only non-negative values, reflecting
-    uncertainty in the variance of match outcomes, with higher values
-    suggesting more variability in outcomes.
+-   `Prior for the Coefficients (year, team2, toss_win)`: Each coefficient is assigned a Normal prior with a mean of 0 and a standard deviation of 2.5, reflecting the assumption that the effects of the predictors are centered around zero with a reasonable amount of uncertainty.
+
+-   `Prior for the Intercept`: A Normal prior with a mean of 0 and standard deviation of 2.5 is used, representing the baseline log-odds of Pakistan winning a match when all predictors are zero.
+
+-   `Prior for the Auxiliary Parameter (prior_aux)`: This represents the dispersion of the logistic regression model and is assigned an Exponential prior with a rate of 1, typical for Bayesian logistic regression.
+It reflects uncertainty in outcome variance, with higher values suggesting more variability.
 
@@
 modelplot(model, coef_omit = "Intercept|team2Ireland") +
-   labs(x = "Coefficient Estimate", y="Predictor Variable")
-```
-
-Analyzing some of these coefficients further, The results from the
-regression model are as follows:
-
--   **Intercept**: The intercept of -55.94 (with a standard error of
-    80.63) represents the log-odds of Pakistan winning a match when all
-    other variables are set to their reference categories. This value is
-    quite large in magnitude, but it is not directly interpretable in
-    isolation. The negative value suggests a very low baseline
-    probability of winning in the absence of other influencing factors.
-
--   **Year**: The coefficient for year is 0.03 (standard error = 0.04).
-    This positive coefficient suggests that, as time progresses, the
-    probability of Pakistan winning a match increases very slightly.
-    However, the small magnitude of the coefficient indicates that this
-    effect is minimal. The relatively narrow standard error suggests
-    that the estimate is reasonably precise, but the practical
-    significance of this trend is modest.
-
--   **Opponent**: The coefficients for the opponent variable represent
-    how the likelihood of Pakistan winning a match changes when facing
-    different teams. These coefficients should be interpreted relative
-    to the reference team, which is not explicitly listed here but is
-    assumed to be the team with the lowest or most neutral effect:
+  labs(x = "Coefficient Estimate", y = "Predictor Variable")
+```
+
+The regression model results are as follows:
+
+-   **Intercept**: The intercept of -55.94 (standard error = 80.63) represents the log-odds of Pakistan winning when all other variables are at their reference categories.
+While large in magnitude, it is not directly interpretable.
+The negative value suggests a very low baseline probability of winning without other influencing factors.
+
+-   **Year**: The coefficient for year is 0.03 (standard error = 0.04), indicating a very slight increase in the probability of Pakistan winning over time.
+The small magnitude and narrow standard error suggest the effect is minimal and the estimate is fairly precise.
+
+-   **Opponent**: The coefficients for opponents show how Pakistan’s likelihood of winning changes against different teams, relative to the reference team:
 
@@
 for (opponent in opponents) {
-
   # Filter the data for the current opponent
@@
     coord_cartesian(ylim = c(0, 14)) +
-    theme_minimal()+
+    theme_minimal() +
     theme(legend.position = "none")
@@
 
-Overall, these graphs provide valuable insights into Pakistan's historical performance trends across different opponents.
+Overall, these graphs provide useful information into Pakistan's historical performance trends across different opponents.
 While some teams, such as Australia and England, have posed significant challenges, others, like Bangladesh and Zimbabwe, have been more favorable matchups for Pakistan.
@@
     winner == "Pakistan" ~ 1,
-    TRUE ~ 0))%>%
+    TRUE ~ 0
+  )) %>%
   mutate(year = lubridate::year(date))
@@
 @fig-career-end-runs-individual presents a scatter plot showing the years in which players retired and the number of runs they scored throughout their careers, with each dot representing an individual Pakistani player.
-The plot reveals that most players retired with fewer than 1,000 career runs, and there is generally no strong positive trend across the data.
+The plot shows that most players retired with fewer than 1,000 career runs, and there is generally no strong positive trend across the data.
 However, there is a notable exception around 2015, where we see a slight uptick in the number of runs.
 This increase is short-lived, as the trend declines again shortly after, until 2024, when a large number of players retired, many of whom had accumulated over 1,000 career runs.
-This scatter plot provides valuable insight into individual player performances, which can be used to explore broader trends in the data.
+This scatter plot provides useful information on individual player performances, which can be used to explore broader trends in the data.
 
@@
 To ensure the analysis focused on relevant information, the data was cleaned and processed to include only Pakistan’s matches, along with important player career statistics that could impact match outcomes.
-By incorporating these factors, the model provided a more comprehensive understanding of Pakistan’s performance dynamics.
-
-The results of the analysis reveal valuable insights into Pakistan's performance trends over time.
+By incorporating these factors, the model provided a more complete understanding of Pakistan’s performance dynamics.
+
+The results of the analysis reveal useful intuitions into Pakistan's performance trends over time.
 For example, the strength of the opposing team played a significant role, with match outcomes showing a clear correlation between Pakistan's success and the competitive level of the opposition.
 Additionally, the toss outcome also emerged as influential factors, providing a deeper understanding of how certain conditions affect Pakistan's likelihood of winning.
-These findings underscore the importance of both internal (team performance) and external (opponent strength, match context) factors in shaping match outcomes, offering strategic insights for future team preparation and performance analysis.
+These findings underscore the importance of both internal (team performance) and external (opponent strength, match context) factors in shaping match outcomes, offering strategic information for future team preparation and performance analysis.
 
@@
 Conversely, the coefficients for teams like India and South Africa point to tougher challenges, with Pakistan's chances of winning being lower in these matchups.
-This pattern highlights the broader strategic landscape, where understanding the relative strength of the opponent becomes crucial for predicting outcomes.
+This pattern highlights the broader strategic landscape, where understanding the relative strength of the opponent becomes important for predicting outcomes.
 It underscores the need for teams to account not just for internal factors, such as player form or match conditions, but also for the competitive level of the opposition when preparing for matches.
 The results validate the intuition that the quality of the opposing team is a significant determinant of Pakistan’s performance, and that adjusting tactics accordingly can improve chances against stronger teams.
-Ultimately, the analysis reinforces the importance of a nuanced, strategic approach to match preparation, where the nature of the opposition is carefully considered alongside team strengths and other match variables.
+Ultimately, the analysis reinforces the importance of a strategic approach to match preparation, where the nature of the opposition is carefully considered alongside team strengths and other match variables.
 
@@
 This can be explained by the presence of a number of veteran players who, with their vast experience, played a pivotal role in guiding the team to more victories.
-Around 2015, several seasoned players reached the peak of their careers or retired, bringing with them not only impressive career statistics but also crucial leadership and expertise on the field.
+Around 2015, several seasoned players reached the peak of their careers or retired, bringing with them not only impressive career statistics but also leadership and expertise on the field.
 The presence of these experienced players undoubtedly helped stabilize the team's performance and contributed to the higher win rates during this period.
@@
 
-Despite the valuable insights gained, there are several limitations in the data and methodology that must be acknowledged.
+Despite the useful information gained, there are several limitations in the data and methodology that must be acknowledged.
 One significant challenge was the limited number of matches involving Pakistan that could be used to train the model.
@@
 For instance, while data on match outcomes (win/loss), toss results, and opponents were available, detailed statistics like the number of runs scored, wickets taken, or batting and bowling performances were only provided for the winning team.
-This lack of comprehensive match data—especially for losing teams—meant that I could not include key performance indicators like runs or wickets in the model, as many of these entries were missing or incomplete.
+This lack of extensive match data—especially for losing teams—meant that I could not include key performance indicators like runs or wickets in the model, as many of these entries were missing or incomplete.
 These missing values could have influenced the accuracy of the model, and it would be prudent to consider these missing data points in future iterations of the analysis, potentially employing imputation methods or seeking alternative data sources.
@@
 - Improving the Data Set: One of the key areas for future research would be to gather a more extensive data set, particularly focusing on adding detailed match statistics, including runs, wickets, and player performances for both teams.
-This would allow for a more comprehensive model, potentially including these factors as predictors in a multivariate analysis.
+This would allow for a more sophisticated model, potentially including these factors as predictors in a multivariate analysis.
 Additionally, expanding the sample size to include more matches across a broader time period could help provide more robust and generalizable results.
@@
 
-- Considering Interaction Effects: Future work could explore more advanced models that account for interactions between variables.
+- Considering Interaction Effects: Future work could explore more sophisticated models that account for interactions between variables.
 For instance, the relationship between Pakistan’s performance and the opponent could vary depending on the specific conditions of a given year, such as the emergence of new players, changes in team strategy, or alterations in the competitive landscape of international cricket.
-Exploring such interaction effects could provide a more nuanced view of Pakistan's performance over time.
+Exploring such interaction effects could provide a more refined view of Pakistan's performance over time.
 
@@
 ## Data Collection
-In an ideal scenario, the data would have been far more comprehensive, particularly with respect to the granular details of each match.
+In an ideal scenario, the data would have been far more extensive, particularly with respect to the granular details of each match.
 To overcome the limitations of the current dataset, it would be essential to gather detailed match data through a combination of observational data and video analysis.
@@
 
-- `Detailed Over-by-Over Data`: Analyzing each over of the match would provide crucial insights into the match's dynamics.
+- `Detailed Over-by-Over Data`: Analyzing each over of the match would provide important data into the match's dynamics.
 This would involve recording key statistics, such as the number of runs scored, the number of wickets taken, the bowler and batter in each over, and any significant events, like boundary hits or wickets.
@@
 This information would allow for a more precise understanding of how the performance of individual players influences the match outcome.
-For instance, if certain players consistently score high or take wickets in critical moments, this could provide valuable insight into the team’s overall performance strategy.
+For instance, if certain players consistently score high or take wickets in key moments, this could provide useful information into the team’s overall performance strategy.
 
@@
 ## Annual Player Performance
-Another critical aspect that could improve the analysis is the collection of annual player performance data.
-While the current dataset allows for some insight into match outcomes, it lacks detailed player statistics across multiple years.
+Another important aspect that could improve the analysis is the collection of annual player performance data.
+While the current dataset allows for some intuition into match outcomes, it lacks detailed player statistics across multiple years.
 To address this, it would be helpful to build a dataset that combines match-level data for each player over the course of several years.
@@
 Key metrics such as the total number of runs scored, the number of wickets taken, the number of matches played, and the number of innings would be compiled.
-This would allow for a more nuanced analysis of individual player performance, considering factors like form and consistency.
+This would allow for a more refined analysis of individual player performance, considering factors like form and consistency.
 
@@
 - `Player and Opponent Comparisons`: To better understand how player performance interacts with match outcomes, it would be useful to collect data on player performances not just for Pakistan, but also for opposing teams.
-By comparing the performances of Pakistan’s players with those of the opposing team’s players, we could gain insights into how the relative strength of players affects the result.
+By comparing the performances of Pakistan’s players with those of the opposing team’s players, we could gain knowledge on how the relative strength of players affects the result.
 For example, if Pakistan’s key players perform well but the opposing team’s star players outperform them, it could explain a loss despite good individual performances.
+
+## Simulation Study
+A key aspect of ensuring the robustness of the data is to evaluate the impact of potential sampling bias on performance predictions.
+By running a simulation study, we could artificially create datasets with varying levels of completeness and bias (e.g., underrepresentation of certain match conditions or players) and test how these biases affect the ability to predict match outcomes.
+For instance, simulations could explore scenarios where certain conditions, such as playing at home or under specific weather conditions, are underrepresented in the data.
+By comparing the outcomes of these simulations with real-world performance predictions, we can better understand how bias in sampling influences model accuracy and the generalizability of results.
+
+## Bias-Adjustment Techniques
+In addition to simulation, statistical techniques such as propensity score matching or inverse probability weighting could be applied to adjust for biases in the observed data.
+These methods would help to correct for any over- or under-representation of certain conditions and ensure that the analysis is reflective of a broader set of match scenarios.
+For example, if the dataset is skewed towards games played against stronger teams, these techniques could help balance the data to ensure that the analysis is not overly influenced by these high-stakes matchups.
 
@@
 performance on unseen data, with the confusion matrix and accuracy
-offering insights into the model’s classification capabilities. While
+offering information on the model’s classification capabilities. While
 accuracy provides a good indication of overall model performance, we
@@
 # Extract relevant fit diagnostics
-fit_diagnostics <- model_summary[c("mean_PPD"),c(1, 3:6)]
+fit_diagnostics <- model_summary[c("mean_PPD"), c(1, 3:6)]
 fit_diagnostics$variable <- "mean_PPD"
@@
   kable(format = "latex")
-
 ```
@@
 model_summary2 <- as.data.frame(model_summary2)
-fit_diagnostics2 <- model_summary2[mcmc_row_names,c(2, 7:8)]
+fit_diagnostics2 <- model_summary2[mcmc_row_names, c(2, 7:8)]
 fit_diagnostics2$variable <- mcmc_row_names

### 2024-12-11 paper/paper.qmd
Update paper.qmd

@@
 date-format: long
-abstract: "This paper analyzes Pakistan's cricket performance from 2002 to 2024, using a linear regression model to examine factors influencing match outcomes.
+abstract: "This paper analyzes Pakistan's cricket performance from 1952 to 2024, using a linear regression model to examine factors influencing match outcomes.
 We find that opponent strength and toss outcomes are significant predictors of Pakistan's success, highlighting both internal team dynamics and external variables.
@@
 Logistic Regression approach, the model incorporates uncertainty in the
-estimates, offers a clear interpretation of the effects of each
+estimates, giving us a clear interpretation of the effects of each
 predictor, and is flexible enough to adapt to new data as it becomes

### 2024-12-11 paper/paper.qmd
Updated typos and added missing reference

@@
 title: "Analyzing Factors Affecting Pakistan Cricket Team's Performance (1952-2024)"
-subtitle: "Opponent and Coin Tossplay Play Key Role in Deciding Outcome"
+subtitle: "Opponent and Coin Toss Play Key Role in Deciding Match Outcome"
 author: 
@@
 date-format: long
-abstract: "This paper analyzes Pakistan's cricket performance from 1952 to 2024, using a linear regression model to examine factors influencing match outcomes.
-We find that opponent strength and toss outcomes are significant predictors of Pakistan's success, highlighting both internal team dynamics and external variables.
+abstract: "This paper analyzes Pakistan's cricket performance from 1952 to 2024, using a Bayesian Logistic Regression model to examine factors influencing match outcomes.
+We find that opponent strength and toss outcomes are significant predictors of Pakistan's success.
 These findings provide information for coaches, analysts, and administrators, offering strategies to improve performance under varying conditions.
@@
 library(modelsummary)
-library(tinytable)
 library(knitr)
@@
 Despite the wealth of player statistics and match highlights available, there has been a lack of systematic statistical analysis that evaluates long-term performance trends.
-This paper fills that gap by analyzing data from 2002 to 2024 to uncover the key factors that contribute to Pakistan’s success in international cricket.
+This paper fills that gap by analyzing data from 1952 to 2024 to uncover the key factors that contribute to Pakistan’s success in international cricket.
 
-We employ a linear regression model to analyze match outcomes,
+We employ a Bayesian Logistic Regression model to analyze match outcomes,
 considering variables like year, opponent, and toss outcomes.
@@
 the reason for selecting the data set we did. @sec-model discusses model
-specification and justification for a Linear Regression model.
+specification and justification for a Bayesian Logistic Regression model.
 @sec-results analyzes the trends and correlations between different
@@
 [@testthat], `caret` [@caret], `rstanarm` [@rstanarm], `tidyr` [@tidyr],
-modelsummary [@modelsummary], tinytable [@tinytable], knitr [@knitr], and ggplot2 [@ggplot2].
+`modelsummary` [@modelsummary], `knitr` [@knitr], `kableExtra` [@kableExtra] and `ggplot2` [@ggplot2].
 
@@
 
-Figures @fig-match-stats show key match statistics for Pakistan.
+@fig-match-stats show key match statistics for Pakistan.
 The data shows that Pakistan frequently faces England, Sri Lanka, and Australia, with a particularly poor record against Australia, including several draws.
@@
 
--   `Prior for the Auxiliary Parameter (prior_aux)`: This represents the dispersion of the logistic regression model and is assigned an Exponential prior with a rate of 1, typical for Bayesian logistic regression.
+-   `Prior for the Auxiliary Parameter (prior_aux)`: This represents the dispersion of the logistic regression model and is assigned an Exponential prior with a rate of 1, typical for Bayesian Logistic Regression.
 It reflects uncertainty in outcome variance, with higher values suggesting more variability.
@@
 
-Our results are summarized in @tbl-model-results, and @fig-model-results
-presents the predicted values of the coefficients. This figure also
+Our results are summarized in @tbl-model-results and @fig-model-results
+presenting the predicted values of the coefficients. The figure also
 includes the 95% confidence intervals for the predictors, as estimated
@@
 
-# table1 <- fit_diagnostics |>
-#   tt()|>
-#   style_tt(j = 2:5, align = "r") |>   # Align columns 2 to 5 to the right
-#   format_tt(digits = 1, num_fmt = "decimal")
-
 # Use kableExtra for table styling
@@
 
-# fit_diagnostics2 |>
-#   tt() |>
-#   style_tt(j = 2:4, align = "r") |>
-#   format_tt(digits = 1, num_fmt = "decimal")
-
 # Use kableExtra for table styling
